### 2015-10-11 Report Aug 5 2015/Report Aug 2015.tex
report  revised until P28

%% bare_jrnl.tex
%% V1.4a
%% 2014/09/17
%% by Michael Shell
%% see http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8a or later) with an IEEE
%% journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_conf_compsoc.tex,
%%                    bare_jrnl_compsoc.tex, bare_jrnl_transmag.tex
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices and paper sizes can       ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

%\documentclass[journal]{IEEEtran}
\documentclass[12pt, draftclsnofoot, onecolumn]{IEEEtran}
%\documentclass[12pt,paper,final]{article}
%
%\documentclass[12pt, onecolummn]{report}
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[journal]{../sty/IEEEtran}

\usepackage[latin1]{inputenc}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{fullpage}
\usepackage{setspace}
\usepackage{graphicx}
%\usepackage[pdftex]{graphicx}
\usepackage{psfrag}
\usepackage{color}
\usepackage{epsfig}
%\usepackage{appendix}
%\usepackage{caption}
\usepackage{cite}
\usepackage{ifpdf}
\usepackage[cmex10]{amsmath}
\usepackage{algorithm}
\usepackage{array}
\usepackage{stfloats}
\usepackage{url}
\usepackage{fixltx2e}
\usepackage{setspace} 
\usepackage{diagbox}
\usepackage{subfigure}
\usepackage{algpseudocode}
\usepackage{multirow}
\usepackage{calc}
% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/tex-archive/info/epslatex/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
\usepackage[cmex10]{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/

% *** FLOAT PACKAGES ***
%
\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Do not use the stfloats baselinefloat ability as IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/dblfloatfix/

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and 
% Axel Sommerfeldt. This package may be useful when used in conjunction with 
% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
% submissions have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.4.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/endfloat/
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/url/
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
%\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
\begin{spacing}{2}
%\doublespacing
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Report: Complex Support Vector Detector for Large MIMO System}

%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%
\author{Tianpei Chen\\
Department of Electrical and Computer Engineering\\
McGill University\\
\today}

%\author{Michael~Shell,~\IEEEmembership{Member,~IEEE,}
%        John~Doe,~\IEEEmembership{Fellow,~OSA,}
%        and~Jane~Doe,~\IEEEmembership{Life~Fellow,~IEEE}% <-this % stops a space
%\thanks{M. Shell is with the Department
%of Electrical and Computer Engineering, Georgia Institute of Technology, Atlanta,
%GA, 30332 USA e-mail: (see http://www.michaelshell.org/contact.html).}% <-this % stops a space
%\thanks{J. Doe and J. Doe are with Anonymous University.}% <-this % stops a space
%\thanks{Manuscript received April 19, 2005; revised September 17, 2014.}}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
%\markboth{Journal of \LaTeX\ Class Files,~Vol.~13, No.~9, September~2014}%
%{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2014 IEEE}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
%\begin{abstract}
%The abstract goes here.
%\end{abstract}

% Note that keywords are not normally used for peerreview papers.
%\begin{IEEEkeywords}
%IEEEtran, journal, \LaTeX, paper, template.
%\end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
%\IEEEPARstart{T}{his} demo file is intended to serve as a ``starter file''
%for IEEE journal papers produced under \LaTeX\ using
%IEEEtran.cls version 1.8a and later.
% You must have at least 2 lines in the paragraph with the drop letter
% (should never be an issue)
%I wish you the best of success.

%\hfill mds

%\hfill September 17, 2014
  One of the biggest challenges the researchers and industry practitioners are facing in wireless communication area is how to bridge the sharp gap between increasing demand of high speed communication of rich multimedia information with high level Quality of Service (QoS) requirements and the limited radio frequency spectrum over a complex space-time varying environment. The most promising technology for solving this problem, Multiple Input Multiple Output (MIMO) technology has been of immense research interest over the last several tens of years is incorporated into the emerging wireless broadband standard like 802.11ac\cite{IEEE802.11ac} long-term evolution (LTE)\cite{3GLTE}.  The core idea of MIMO system is to use multiple antennas at both transmitting and receiving end, so that multiplexing gain (multiple parallel spatial data pipelines that can improve bandwidth efficiency) and diversity gain (better reliability of communication link) is obtained by exploiting the spatial domain. Large MIMO (also called Massive MIMO) is an upgraded version of conventional MIMO technology employing hundreds of low power low price antennas at base station (BS), that serves several tens of terminals simultaneously. This technology can achieve full potential of conventional MIMO system while providing additional power efficiency as well as system robustness both to unintended man-made interference and intentional jamming.\cite{rusek2013scaling}\cite{larsson2014massive}.

  The price paid for large MIMO system is the increased complexities for signal processing at both transmitting and receiving end. The uplink Detector is one of the key components in large MIMO systems. With orders magnitude more antennas at the BS, benefits and challenges coexist in designing of detection algorithms for the uplink communication of large MIMO systems. On one hand, a large number of receive antennas provide potential of large diversity gains, on the other hand, complexity of the algorithm becomes crucial to make the system practical.

Vertical Bell Laboratories Layered Space-Time (V-BLAST) architecture for MIMO system can achieve high spectrum efficiency by spatial multiplexing (SM), that is, each transmit antenna transmits independent symbol streams. However the optimal maximum likelihood detector (MLD) for V-BLAST systems that perform exhaustive search has a complexity that increases exponentially with number of transmitted antennas, which is prohibitive for practical applications.

As alternatives to MLD, linear detectors (LD) such as zero-forcing (ZF) and minimum mean square error (MMSE) with optimized ordering sequential interference cancellation (ZF-OSIC, MMSE-OSIC) are exploited in V-BLAST architecture\cite{wolniansky1998v}\cite{foschini1999simplified}\cite{benesty2003fast},  however the performance of ZF-OSIC and MMMSE-OSIC are inferior comparing to MLD.

Sphere Decoder (SD)\cite{damen2003maximum} is the most prominent algorithm that utilizes the lattice structure of MIMO systems, which can achieve optimal performance with relatively much lower complexity comparing to MLD. However, SD has two major shortages that make it problematic to be integrated into a practical systems. The first shortage is SD has various complexities under different signal to noise ratios (SNR), while a constant processing data rate is required for hardware. The second shortage is SD's complexity still has a lower bound for complexity that increases exponentially with the number of transmit antennas and the order of modulation scheme\cite{jalden2005complexity}.
The fixed complexity sphere decoder (FCSD)\cite{barbero2008fixing} makes it possible to achieve near optimal performance with a fixed complexity under different value of SNR. The FCSD inherits the principle of list based searching algorithms, which first generate a list of candidate symbol vectors and then the best candidate is chosen as the solution. The other sub optimal detectors belong to this class include Generalized Parallel Interference Cancellation (GPIC)\cite{luo2008generalized} and Selection based MMSE-OSIC(sel-MMSE-OSIC)\cite{radji2009interference}. However, all these list based searching algorithms have the same shortage - their complexities increase exponentially with the number of transmit antennas and the order of modulation scheme\cite{radji2009interference}. Therefore, such algorithms are prohibitive when it comes to a large number of antennas or a high order modulation scheme, for example in IEEE 802.11ac standard\cite{IEEE802.11ac}, the modulation scheme is 256QAM.

%Linear detectors (LD) like minimum mean square error (MMSE) and zero forcing (ZF) along with their sequential interference cancellation with optimized ordering (OSIC) counterparts\cite{wolniansky1998v}, which have good performance for low loading factor in massive MIMO system (the number of receive antennas is much larger than the number of transmit antennas)\cite{hoydis2013massive}. 
Besides the above detection algorithms designed for conventional MIMO systems, in the last several years, a set of detection algorithms have been proposed for large MIMO systems with complexities that are comparable with MMSE detector and near-optimal performance. such algorithms include likelihood ascend searching (LAS) algorithms\cite{vardhan2008low}\cite{li2010multiple}, %An theoretical analysis of upper bound of bit error rate (BER) and lower bound of on asymptotic multiuser efficiency for the LAS detector was also presented\cite{sunfamily}.
Tabu search based algorithms which have superior performance compared to LAS detectors because local minima can be avoided (e.g. Layered Tabu search (LTS)\cite{srinidhi2011layered}, Random Restart Reactive Tabu search (R3TS)\cite{datta2010random}), Message passing technique based algorithms (e.g. Belief propagation (BP) detectors based on graphic model and Gaussian Approximation (GA)\cite{som2011low}\cite{som2010improved}\cite{narasimhan2014channel}\cite{goldberger2011mimo}), Probabilistic Data Association based algorithms \cite{mohammed2009low}, Monte Carlo sampling based algorithms (e.g. Markov Chain Monte Carlo (MCMC) algorithm\cite{datta2013novel}) and Lattice Reduction (LR) aided algorithms\cite{zhou2013element}.

Firmly grounded in framework of statistical learning theory, the Support Vector Machine (SVM) technique has become a powerful tool to solve real world supervised learning problems such as classification, regression and prediction. the SVM method is a nonlinear generalization of Generalized Portrait algorithm developed by Vapnik in 1960s\cite{vapnik1963pattern}\cite{vapnik1964note}, which can provide good generalization performance\cite{scholkopf2002learning}.

Interest in SVM boosted since 1990s, promoted by the works of Vapnik and co-workers at AT\& T Bell laboratory\cite{boser1992training}\cite{guyon1993automatic}\cite{vapnik2013nature}\cite{cortes1995support}\cite{scholkopf1996incorporating}\cite{vapnik1996support}.
Moreover, the kernel based methods\cite{scholkopf2002learning} solve nonlinear learning tasks by mapping input data sets into high dimensional feature spaces, and replacing inner products of feature mappings by computational inexpensive kernel functions discarding the actual structure of the feature space. This rationale is supported mathematically by the notion of Reproducing Kernel Hilbert Space (RKHS). 
Based on the same regularized risk function principle, $\epsilon$-Support Vector Regression ($\epsilon$-SVR) was developed\cite{vapnik2013nature}\cite{smola2004tutorial}.

Similar to SVM, the $\epsilon$-SVR solves an original optimization problem by transforming it into a Lagrange dual optimization problem, which can be solved by Quadratic Programming (QP). Sequential Minimal Optimization (SMO) algorithm was proposed as a fast algorithm to solve this QP problem by decomposing the it into sub QP problems and solving them analytically\cite{platt1999fast}. Therefore, the computational intensive numerical method can be avoided. A more general method is decomposition solver, which refers to a set of algorithms that separate the optimization variables (Lagrange multipliers) into two sets W and N, W is the work set and N contains the remaining optimization variables. In each iteration, only the optimization variables in the work set is optimized while keeping other variables fixed. The SMO algorithm is an extreme case of decomposition solver. An important issue of decomposition solver is the choice of the work set. One strategy is to choose Karush-Kuhn-Tucker (KKT) condition violators, ensuring final converge\cite{osuna1997improved}. Because of the linear constraint inducted by offset, the SMO algorithm restricts the size of the work set to 2. In\cite{steinwart2011training}, a method to train SVM without offset was proposed, with the comparable performance to the SVM with offset. A set of sequential single variable work set selection strategies, which require $O(n)$ searching time are proposed. The optimal double variable work set selection strategy, which performs exhaustively searching, however, requires $O(n^{2})$ searching time. The authors demonstrate that with the combination of two proposed single variable work set selection strategies, convergence can be achieved by a iteration time that is as few as optimal double variable work set selection strategy.

The mathematical foundation of kernel based methods is RKHS which is defined in complex domain, however most of the practitioners are dealing with real data sets. In communication and signal processing area, the channel gains, signals, waveforms etc. are all represented in complex form. Recently, a pure complex SVR \& SVM based on complex kernel was proposed in\cite{bouboulis2013complex}, which can deal with the complex data set purely in complex domain. The results in\cite{bouboulis2013complex} demonstrate better performances as well as reduced complexity comparing to simply split learning task into two real case by real kernel.  
Based on this work, we derive of a complexity-performance controllable detector for large MIMO systems based on a dual channel complex SVR (CSVR). The detector can work in two parallel real SVR channels which can be solved independently. Moreover, only the real part of kernel matrix is needed in both channels. This means a large amount of computation can be reduced.
Based on the discrete time MIMO channel model, in our regression model, this CSVR-detector
is constructed without offset, Therefore, for each real SVR without offset, in principle, only one variable is needed to be updated in each iteration, In our scheme, a sequential single variable selection strategy is proposed. By this strategy, two variables can be updated at each iteration, with much smaller searching time.

%\subsection{Subsection Heading Here}
%Subsection text here.

% needed in second column of first page if using \IEEEpubid
%\IEEEpubidadjcol

%\subsubsection{Subsubsection Heading Here}
%Subsubsection text here.

\section{Brief Introduction to $\epsilon$-Support Vector Regression}\label{Introduce epsilon SVR}
\subsection{Regression Model}
Suppose we are given training data set $((\mathbf{x}_{1}, y_{1}),(\mathbf{x}_{2},y_{2}),\cdots,(\mathbf{x}_{L},y_{L}))$, $L$ denotes the number of training samples, $\mathbf{x}_{i}\in \mathbb{R}^{V}$ denotes input data vector, $V$ is the number of features in $\mathbf{x}_{i}$. $y_{i}$ denotes output. Let $\mathbf{w}$ denotes regression coefficient vector, $\Phi(\mathbf{x}_{i})$ denotes the mapping of $\mathbf{x}_{i}$ to higher dimensional feature space, $\mathbf{w},\Phi(\mathbf{x}_{i})\in \mathbb{R}^{\Lambda}$, $\Lambda \in \mathbb{R}$ denotes the dimension of mapped feature space (For linear model, $\mathbf{x}_{i}=\Phi(\mathbf{x}_{i})$). The regression model (either linear or non-linear) is given by 
\begin{equation}
y_{i}=\mathbf{w}^{T}\Phi(\mathbf{x}_{i})+b  \quad i= 1\cdots L 
\label{equation1}
\end{equation}

\begin{figure}
\centering
\def\svgwidth{\columnwidth}
\input{epsilon-SVR.pdf_tex}
\caption{$\epsilon$-Support Vector Regression and Risk Functional}
\label{epsilon-SVR}
\end{figure}

We present the primal optimization problem directly
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
\min_{\mathbf{w}} f(\mathbf{w})=\frac{1}{2}||\mathbf{w}||^{2}+C\sum_{i=1}^{L}(R(\xi_{i})+R(\hat{\xi}_{i}))\\
s.t. \left\{\begin{array}{ll}
y_{i}-\mathbf{w}^{T}\Phi(\mathbf{x}_{i})-b \leq \epsilon+\xi_{i}, i=1,2\cdots, L \\
\mathbf{w}^{T}\Phi(\mathbf{x}_{i})+b-y_{i} \leq \epsilon+\hat{\xi}_{i}, i=1,2\cdots, L\\
\epsilon_{i}, \xi_{i},\hat{\xi}_{i} \geq 0, i=1,2\cdots, L
\end{array}\right.
\label{primal optimization problem}
\end{IEEEeqnarray}
In (\ref{primal optimization problem}), $\frac{1}{2}||\mathbf{w}||^{2}$ is the regularization term in order to ensure the flatness of regression model, $\epsilon$ denotes the precision, As shown in Fig \ref{epsilon-SVR}, the area between two dash line is called $\epsilon$ tube. Only those data samples located outside $\epsilon$ tube contribute to cost. Furthermore $\xi_{i}$ and $\hat{\xi}_{i}$ denote slack variables that cope with noise of input data samples, $R(u)$ denotes risk function, the simplest risk function is $R(u)=u$, The type of cost function is determined by the statistical distribution of noise\cite{smola2004tutorial}. For example if the noise is Gaussian noise, then the optimal cost function is $R(u)=\frac{1}{2}u^{2}$. The term $C\sum_{i=1}^{L}(R(\xi_{i})+R(\hat{\xi}_{i}))$ denotes the penalty of noise, $C\in \mathbb{R}$ and $C\geq 0$ that controls the trade off between regularization term and cost function term.

\subsection{Cost Function}\label{cost function}
The optimal cost function in (\ref{primal optimization problem}) can be derived based on maximum likelihood (ML) principle. Assume the data samples $\mathbf{x}_{i}$ in data set are iid, Let $f_{true}(\mathbf{x}_{i}), i=1,2,\cdots L$ denotes true regression function. the underlying assumption is $y_{i}=f_{true}(\mathbf{x}_{i})+\xi_{i}, i=1,2,\cdots, L$, $\xi_{i}$ denotes additive noise of the $i$th data sample, with probability density function (pdf) $Pr(\cdot)$. Let $P(\cdot)$ denotes the pdf of $y_{i}$. Based on ML principle we want to 
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
\max_{f}\quad \prod_{i=1}^{L}P(y_{i}|f(\mathbf{x}_{i}))=\prod_{i=1}^{L}P(f(\mathbf{x}_{i})+\xi_{i}|f(\mathbf{x}_{i}))=\prod_{i=1}^{L}Pr(\xi_{i})=\prod_{i=1}^{L}Pr(y_{i}-f(\mathbf{x}_{i}))\\
\label{cost function1}
\end{IEEEeqnarray}
Take the logarithm of $\prod_{i=1}^{L}Pr(y_{i}-f(\mathbf{x}_{i}))$, we have
\begin{IEEEeqnarray}[\relax]{l}
\sum_{i=1}^{L}log(Pr(y_{i}-f(\mathbf{x}_{i}))),
\label{cost function2}
\end{IEEEeqnarray}
maximizing (\ref{cost function2}) is equivalent to minimizing $-\sum_{i=1}^{L}log(Pr(y_{i}-f(\mathbf{x}_{i})))$
Let $c(\mathbf{x}_{i}, y_{i}, f(\mathbf{x}_{i}))$ denotes the $i$th cost function, which is defined as  
\begin{IEEEeqnarray}[\relax]{l}
c(\mathbf{x}_{i}, y_{i}, f(\mathbf{x}_{i}))=-log(P(y_{i}-f(\mathbf{x}_{i}))).
\label{cost function3}
\end{IEEEeqnarray}
Thus (\ref{cost function1}) can be rewritten as 
\begin{equation}
\min_{f}\quad  \sum_{i=1}^{L}c(\mathbf{x}_{i}, y_{i}, f(\mathbf{x}_{i})),
\label{Total risk function}
\end{equation} 
In $\epsilon$-SVR, in order to provide more flexibility to precision control, Vapnik's $\epsilon$-insensitive function, as shown in (\ref{Vepsilon}), is applied to (\ref{cost function3}).
\begin{IEEEeqnarray}[\relax]{l}
|u|_{\epsilon}=\left\{\begin{array}{ll}
0   &if\quad |u|<\epsilon\\
|u|-\epsilon  &otherwise\\
\end{array}\right.
\label{Vepsilon}
\end{IEEEeqnarray}
Thus the final form of cost function in $\epsilon$-SVR can be written as 
\begin{equation}
\tilde{c}(y_{i}, \mathbf{x}_{i}, f(\mathbf{x}_{i}))=-log(Pr(|y_{i}-f(\mathbf{x}_{i})|_{\epsilon})),
\label{final cost function}
\end{equation}
Consider the cost function term in (\ref{primal optimization problem}),
\begin{equation}
\sum_{i=1}^{L}R(\xi_{i})+R(\hat{\xi}_{i})=\sum_{i=1}^{L}\tilde{c}(y_{i}, \mathbf{x}_{i}, f(\mathbf{x}_{i}))=\sum_{i=1}^{L}-log(Pr(|y_{i}-f(\mathbf{x}_{i})|_{\epsilon}))
\end{equation}  
the cost function term is determined according to noise distribution.  
%where $m_{i}\in \mathbb{R}$, $m_{i}>0$ denotes the weight parameter, if $y_{i}>f_{estiamtion}(\mathbf{x})$, $m_{i}=m_{positive}$, else $m_{i}=m_{negative}$, Therefore the regularized risk function can written as 
%\begin{equation}
%minimize\quad \lambda||\mathbf{w}||^{2}+\tilde{c}(\mathbf{x}, y, f(\mathbf{x})),
%\label{total cost function}
%\end{equation} 
%where $\lambda$ denotes the weight of regularization term, divide (\ref{total cost function}) by $\frac{1}{2\lambda}$, we have the optimization problem  
%\begin{equation}
%minimize \quad \frac{1}{2}||\mathbf{w}||^{2}+\sum_{i=1}^{l}C_{i}(-log(P(|y_{i}-f(\mathbf{x}_{i})|_{\epsilon}))),
%\label{final cost function}
%\end{equation}
%where $C_{i}=\frac{m_{i}}{2\lambda l}$, based on (\ref{final cost function}), by introducing slack variables, we can easily derive the equivalent optimization problem as same as (\ref{primal optimization problem}):
%\begin{eqnarray}
%\nonumber
%minimize \quad f(\mathbf{w})=\frac{1}{2}||\mathbf{w}||^{2}+\sum_{j=1}^{l}C_{i}(R(\xi_{i})+R(\hat{\xi}_{i}))\\
%s.t. \left\{\begin{array}{ll}
%y_{i}-\mathbf{w}^{T}\Phi(\mathbf{x}_{i})-b &\leq \epsilon+\xi_{i}\\
%\mathbf{w}^{T}\Phi(\mathbf{x}_{i})+b-y_{i} &\leq \epsilon+\hat{\xi}_{i}\\
%\epsilon, \xi,\hat{\xi} &\geq 0\\
%\end{array}\right.
%\label{primal objective function}
%\end{eqnarray}
%where $R(x)=-log(P(x))$, by this way, the discontinuity of $\epsilon$-insensitive function is conquered. 
\subsection{Lagrange Duality}\label{section lagrange duality}
According to Lagrange Theorem, the constraint optimization problem (\ref{primal optimization problem}) can be transformed to Lagrangian dual form by combining the original optimization function with inequality constraints, the combination coefficient is called Lagrange multiplier. The Lagrange function is given by. 
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
L(\mathbf{w}, b, \xi_{i}, \hat{\xi}_{i}, \alpha_{i}, \hat{\alpha}_{i}, \eta_{i}, \hat{\eta}_{i})=
\frac{1}{2}||\mathbf{w}||^{2}+C\sum_{i=1}^{L}(R(\xi_{i})+R(\hat{\xi}_{i}))-\sum_{i=1}^{L}(\eta_{i}\xi_{i}+\hat{\eta}_{i}\hat{\xi}_{i})\\
\nonumber
+\sum_{i=1}^{L}\alpha_{i}(y_{i}-\mathbf{w}^{T}\Phi(\mathbf{x}_{i})-b-\epsilon-\xi_{i})+\sum_{i=1}^{L}\hat{\alpha}_{i}(\mathbf{w}^{T}\Phi(\mathbf{x}_{i})+b-y_{i}-\epsilon-\hat{\xi}_{i})\\
s.t. \left\{\begin{array}{cc}
\eta_{i}, \hat{\eta}_{i}, \alpha_{i}, \hat{\alpha}_{i}\geq 0, i=1,2,\cdots L\\
\xi_{i}, \hat{\xi}_{i}\geq  0, i=1,2,\cdots L\\
\end{array}\right.
\label{lagrange duality1}
\end{IEEEeqnarray}
where $\eta_{i}$, $\hat{\eta}_{i}$, $\alpha_{i}$, $\hat{\alpha}_{i}$ are Lagrange multipliers.

The sufficient and necessary conditions such that a solution $\mathbf{w}$ of the constrained optimization problem in (\ref{primal optimization problem}) satisfies, are called Karush-Kuhn-Tucker (KKT) conditions. Here we elaborate a little further about how the dual objective problem is derived from KKT conditions.

Assume a constraint optimization problem is given by 
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
\min_{\mathbf{w}} \quad f(\mathbf{w})\\
s.t. \quad c_{i}(\mathbf{w})\leq 0, i=1,2,\ldots L,
\label{constraint optimization problem}
\end{IEEEeqnarray}
its Lagrange function is given by 
\begin{equation}
L(\mathbf{w}, a_{i})=f(\mathbf{w})+\sum_{i=1}^{L}a_{i}c_{i}(\mathbf{w}),
\label{orginal Lagrange}
\end{equation}
where $a_{i}$ denote Lagrange multipliers.
 Based on Theorem 6.21 in \cite{scholkopf2002learning}, For a variable set $[\bar{\mathbf{w}}, \bar{a}_{i}]$, $\bar{\mathbf{w}}$ is the solution to (\ref{constraint optimization problem}) only when the following inequalities are satisfied
 \begin{equation}
 L(\mathbf{w}, \bar{a}_{i})\geq L(\bar{\mathbf{w}}, \bar{a}_{i})\geq L(\bar{\mathbf{w}}, a_{i})
 \label{KKT inequalities}
 \end{equation}
This inequalities yield KKT conditions (see Theorem 6.26 \cite{scholkopf2002learning}), which are 
\begin{IEEEeqnarray}[\relax]{l}
\label{original KKT condition1}
\partial_{\mathbf{w}}L(\bar{\mathbf{w}}, a_{i})=\partial_{\mathbf{w}}f(\bar{\mathbf{w}})+\sum_{i=1}^{L}a_{i}\partial_{\mathbf{w}}c_{i}(\bar{\mathbf{w}})=0,\\
\label{original KKT condition2}
\partial_{a_{i}}L(\bar{\mathbf{w}}, \bar{a}_{i})=c_{i}(\bar{\mathbf{w}})\leq 0, i=1,2,\ldots L\\
\label{original KKT condition3}
\bar{a}_{i}c_{i}(\bar{\mathbf{w}})=0, i=1,2,\ldots L
\end{IEEEeqnarray}
In order to satisfy the first inequality in (\ref{KKT inequalities}), (\ref{original KKT condition1}) has to hold,
applying (\ref{original KKT condition1}) to (\ref{lagrange duality1}), which are  
\begin{IEEEeqnarray}[\relax]{l}
\label{partial1}
\frac{\partial L}{\partial \mathbf{w}}=\mathbf{w}-\sum_{i=1}^{l}(\alpha_{i}-\hat{\alpha}_{i})\Phi(\mathbf{x}_{i})=0\\
\label{partial2}
\frac{\partial L}{\partial \xi_{i}}=C_{i}R^{'}(\xi_{i})-\eta_{i}-\alpha_{i}=0, i=1,2,\cdots L\\
\label{parial3}
\frac{\partial L}{\partial \hat{\xi}_{i}}=C_{i}R^{'}(\hat{\xi_{i}})-\hat{\eta}_{i}-\hat{\alpha}_{i}=0, i=1,2,\cdots L\\
\label{partial4}
\frac{\partial L}{\partial b}=\sum_{i=1}^{l}(\alpha_{i}-\hat{\alpha}_{i})=0\\\nonumber
\end{IEEEeqnarray}
Then by substituting (\ref{partial1})-(\ref{partial4}) to (\ref{lagrange duality1}), (\ref{lagrange duality1}) can be rewritten as :
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
\theta(\alpha_{i}, \hat{\alpha}_{i})=\frac{1}{2}\sum_{i=1}^{L}\sum_{j=1}^{L}(\alpha_{i}-\hat{\alpha_{i}})(\alpha_{j}-\hat{\alpha_{j}})\Phi^{T}(\mathbf{x}_{j})\Phi(\mathbf{x}_{i})+C\sum_{i=1}^{L}[R(\xi_{i})+R(\hat{\xi}_{i})]-\sum_{i=1}^{L}[(CR^{'}(\xi_{i})-\alpha_{i})\xi_{i}\\
\nonumber
+(CR^{'}(\hat{\xi}_{i})-\hat{\alpha}_{i})\hat{\xi}_{i}]+\sum_{i=1}^{L}\alpha_{i}(y_{i}-\sum_{j=1}^{L}(\alpha_{j}-\hat{\alpha}_{j})\Phi^{T}(\mathbf{x}_{j})\Phi(\mathbf{x}_{i})-b-\epsilon-\xi_{i})+\\
\sum_{i=1}^{L}\hat{\alpha}_{i}(\sum_{j=1}^{L}(\alpha_{j}-\hat{\alpha}_{j})\Phi^{T}(\mathbf{x}_{j})\Phi(\mathbf{x}_{i})+b-y_{i}-\epsilon-\hat{\xi}_{i})
\label{lagrange duality1 inter}
\end{IEEEeqnarray}
notice in (\ref{partial4}), $\sum_{i=1}^{L}(\alpha_{i}-\hat{\alpha}_{i})=0$, define $\tilde{R}(u)=R(u)-uR^{'}(u)$, (\ref{lagrange duality1 inter}) can be further simplified to
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
\theta(\alpha_{i}, \hat{\alpha}_{i}) =-\frac{1}{2}\sum_{i=1}^{L}\sum_{i=1}^{L}(\alpha_{i}-\hat{\alpha_{i}})(\alpha_{j}-\hat{\alpha_{j}})\Phi(\mathbf{x}_{j})^{T}\Phi(\mathbf{x}_{i})+C\sum_{i=1}^{L}[(\tilde{R}(\xi_{i})+\tilde{R}(\hat{\xi}_{i})]\\
\nonumber
+\sum_{i=1}^{L}[(\alpha_{i}-\hat{\alpha_{i}})y_{i}-(\alpha_{i}+\hat{\alpha_{i}})\epsilon]\\
s.t.\left\{\begin{array}{l}
\sum_{i=1}^{l}(\alpha_{i}-\hat{\alpha_{i}})=0\\
0<\alpha<C\tilde{R}^{'}(\alpha)\\
0<\hat{\alpha}<C\tilde{R}^{'}(\hat{\alpha})\\
\end{array}\right.
\label{dual objective function1}
\end{IEEEeqnarray}
In order to satisfy the second inequality in (\ref{KKT inequalities}), (\ref{original KKT condition2}) and  (\ref{original KKT condition3}) have to hold. Condition (\ref{original KKT condition2}) is satisfied when the maximum of $L(\bar{\mathbf{w}}, a_{i})$ is found, notice $\theta(\alpha_{i}, \hat{\alpha}_{i})$ is equivalent to $L(\bar{\mathbf{w}}, a_{i})$ in (\ref{KKT inequalities}), thus yielding the dual optimization problem, which is given by 
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
\max_{\alpha_{i}, \hat{\alpha}_{i}}\quad \theta(\alpha_{i}, \hat{\alpha}_{i}) =-\frac{1}{2}\sum_{i=1}^{L}\sum_{j=1}^{L}(\alpha_{i}-\hat{\alpha_{i}})(\alpha_{j}-\hat{\alpha_{j}})\Phi^{T}(\mathbf{x}_{j})\Phi(\mathbf{x}_{i})+C\sum_{i=1}^{L}(\tilde{R}(\xi_{i})+\tilde{R}(\hat{\xi}_{i}))\\
+\sum_{i=i}^{L}[(\alpha_{i}-\hat{\alpha_{i}})y_{i}-(\alpha_{i}+\hat{\alpha_{i}})\epsilon]
\label{dual optimization function2}
\end{IEEEeqnarray}
Define $\mathbf{a}=[\alpha_{1},\alpha_{2}, \ldots, \alpha_{L}]^{T} $, $\mathbf{\hat{a}}=[\hat{\alpha}_{1},\hat{\alpha}_{2}, \ldots, \hat{\alpha}_{L}]^{T} $, $\mathbf{y}=[y_{1}, y_{2}, \ldots y_{L}]^{T}$, $ \mathbf{e}=[1,1,\ldots,1]^{T}\in \mathbb{R}^{L}$, $\mathbf{e}_{i}$ denotes the vector that only $i$th component is 1 while the rest are all 0, $\mathbf{R}_{\xi}=[\tilde{R}(\xi_{1}), \tilde{R}(\xi_{2}), \ldots, \tilde{R}(\xi_{L})]^{T}$, $\mathbf{R}_{\hat{\xi}}=[\tilde{R}(\hat{\xi}_{1}), \tilde{R}(\hat{\xi}_{2}), \ldots, \tilde{R}(\hat{\xi}_{L})]^{T}$, $[\mathbf{K}]_{ij}=\Phi(\mathbf{x}_{j})^{T}\Phi(\mathbf{x}_{i})$ denotes a component of data kernel matrix at $i$th row and $j$th column. An alternative vector form of (\ref{dual optimization function2}) can be written as 
\begin{equation}
\max_{\mathbf{a}, \hat{\mathbf{a}}}\theta(\mathbf{a},\hat{\mathbf{a}})=-\frac{1}{2}(\mathbf{a}-\mathbf{\hat{a}})^{T}\mathbf{K}(\mathbf{a}-\mathbf{\hat{a}})+(\mathbf{y}-\epsilon \mathbf{e})^{T}\mathbf{a}+(-\mathbf{y}-\epsilon\mathbf{e})^{T}\mathbf{\hat{a}}+C\mathbf{e}^{T}(\mathbf{R}_{\xi}+\mathbf{R}_{\hat{\xi}}),
\label{dual objective alternative}
\end{equation} 
We define the following $2L$ vectors 
$\mathbf{a}^{(*)}$= [$\mathbf{a} \atop \mathbf{\hat{a}}$], $\mathbf{v}\in \mathbb{R}^{2L}$, 
\begin{IEEEeqnarray}[\relax]{l}
[\mathbf{v}]_{i}=\left\{\begin{array}{ll}
1\quad i=1,\ldots, l\\
-1\quad i=l+1,\ldots, 2l\\ 
\end{array}\right.
\end{IEEEeqnarray}
(\ref{dual objective alternative}) can also be reformulate as 
\newcommand{\mysmallarraydecl}{\renewcommand{%
\IEEEeqnarraymathstyle}{\scriptscriptstyle}%
\renewcommand{\IEEEeqnarraytextstyle}{\scriptsize}%
\renewcommand{\baselinestretch}{1.1}%
\settowidth{\normalbaselineskip}{\scriptsize
\hspace{\baselinestretch\baselineskip}}%
\setlength{\baselineskip}{\normalbaselineskip}%
\setlength{\jot}{0.25\normalbaselineskip}%
\setlength{\arraycolsep}{2pt}}
\begin{equation}
\max_{\mathbf{a}^{*}}\quad \theta(\mathbf{a}^{*})=-\frac{1}{2}(\mathbf{a}^{*})^{T}\left[\begin{IEEEeqnarraybox*}[\mysmallarraydecl][c]{,c/c,}
\mathbf{K}& -\mathbf{K}\\
-\mathbf{K}& \mathbf{K}
\end{IEEEeqnarraybox*}\right]
\mathbf{a}^{(*)}+[(\mathbf{y}-\epsilon)^{T}, (-\mathbf{y}-\epsilon)^{T}]\mathbf{a}^{(*)}+C\mathbf{e}^{T}(\mathbf{R}_{\xi}+\mathbf{R}_{\hat{\xi}}),
\label{vectorize dual optimization function}
\end{equation}

Condition in (\ref{original KKT condition3}) is called KKT complementary condition, the value of $\sum_{i=1}^{L}\bar{a}_{i}c_{i}(\bar{\mathbf{w}})$ can be used to monitor how close the solution is to the global optimum. Thus it can be used as a stopping criterion. In the constraint optimization problem of (\ref{lagrange duality1}), the KKT complementary conditions are given by 
\begin{IEEEeqnarray}[\relax]{l}
\left\{\begin{array}{cc}
\alpha_{i}(y_{i}-\mathbf{w}^{T}\Phi(\mathbf{x}_{i})-b-\epsilon-\xi_{i})=0, i=1,2\cdots L\\
\hat{\alpha}_{i}(\mathbf{w}^{T}\Phi(\mathbf{x}_{i})+b-y_{i}-\epsilon-\hat{\xi}_{i})=0, i=1,2\cdots L
\end{array}\right. 
\label{KKT complimentary}
\end{IEEEeqnarray}

%The conditions in (\ref{KKT complimentary}) are called KKT complimentary condition. 
%Obviously by taking partial derivations of (\ref{lagrange duality1}) with respect to $\mathbf{w}$, $\xi_{i}$, $\hat{\xi}_{i}$ and $b$, $\theta(\alpha_{i},\hat{\alpha}_{i})\leq \Theta(\mathbf{w},\alpha_{i}, \hat{\alpha}_{i}, \eta_{i}, \hat{\eta}_{i})$, notice in $\theta(\alpha_{i}, \hat{\alpha}_{i})$ KKT conditions in (\ref{partial1}) to (\ref{partial4}) are satisfied. Therefore, KKT complimentary conditions (\ref{KKT complimentary}) is the only requirement to find global optimal point of original optimization problem. Because 
%\begin{equation}
%\Theta(\mathbf{w},\alpha, \hat{\alpha}, \eta, \hat{\eta})=f(\mathbf{w})+\sum_{i}(\alpha_{i} g_{i}(\mathbf{w})+\hat{\alpha}_{i}\hat{g}_{i}(\mathbf{w}))+\sum_{i}(\eta_{i} l_{i}(\mathbf{w})+\hat{\eta}_{i}\hat{l}_{i}(\mathbf{w})),\label{Theta function}
%\end{equation} 
%where $g_{i}(\mathbf{w})$, $\hat{g_{i}(\mathbf{w})}$, $l_{i}(\mathbf{w})$, $\hat{l}_{i}(\mathbf{w})$ denote inequality constraints.
%\begin{IEEEeqnarray}[\relax]{l}
%g_{i}(\mathbf{w})=\mathbf{y}_{i}-\mathbf{w}^{T}\Phi(\mathbf{x}_{i})-b-\epsilon-\xi_{i}\leq 0\\
%\hat{g}_{i}(\mathbf{w})=\mathbf{w}^{T}\Phi(\mathbf{x}_{i})+b-\mathbf{y}_{i}-\epsilon-\hat{\xi}_{i}\leq 0\\
%l_{i}(\mathbf{w})=-\xi_{i}\leq 0\\
%\hat{l}_{i}(\mathbf{w})=-\hat{\xi}_{i}\leq 0\\
%\nonumber
%\end{IEEEeqnarray}
%Hence $\Theta\leq f(\mathbf{w})$, we have $\theta\leq \Theta\leq f(\mathbf{w})$, therefore the upper bound of $\theta$ is determined by the original objective function $f(\mathbf{w})$. when $\theta=f(\mathbf{w})$,  according to (\ref{Theta function}), the linear combination term of inequality constraints equal to zero, that is, KKT complimentary conditions in (\ref{KKT complimentary}) satisfied. hence the global optimal point is found for both $\theta$ and $f(\mathbf{w})$ if and only if the equality holds. Therefore duality gap is defined as $G=f(\mathbf{w})-\theta$, which can be used as an evaluation for the closeness of one solution to the global optimal.

%In conclusion, the dual objective function can be written as

\section{Dual Channel Complex Support Vector detection for Large MIMO system}\label{dual channel CSVR}
\subsection{System Model}\label{system model}
  Consider a uncoded complex large MIMO uplink spatial multiplexing (SM) system with $N_{t}$ users, where each is equipped with transmit antenna. The number of receive antennas at the Base Station (BS) is $N_{r}$, $N_{r}\geq N_{t}$. Typically large MIMO systems have hundreds of antennas at the BS, as shown in Fig {\ref{large MIMO uplink model}}.
  \begin{figure}[htb]
  \centering
  \def\svgwidth{\columnwidth}
  %\includegraphics[scale=•]{•}
  \input{largeMIMOuplink.pdf_tex}
  \caption{Large MIMO uplink system }
  \label{large MIMO uplink model}
  \end{figure}

Bit sequences, which are modulated to complex symbols, are transmitted by the users over a flat fading channel. The discrete time model of the system is given by:
\begin{equation}
\mathbf{y}=\mathbf{H}\mathbf{s}+\mathbf{n},   \label{discrete time MIMO system}
\end{equation}
where $\mathbf{y}\in\mathbb{C}^{N_{r}\times 1}$ is the received symbol vector, $\mathbf{s}\in \mathbb{C}^{N_{t}}$ is the transmitted symbol vector, with components that are mutually independent and taken from a finite signal constellation alphabet $\mathbb{O}$ (e.g. BPSK, 4-QAM, 16-QAM, 64-QAM), $|\mathbb{O}|=M$. The transmitted symbol vectors $\mathbf{s}\in \mathbb{O}^{N_{t}}$, satisfy $\mathbb{E}[\mathbf{s}\mathbf{s}^{H}]=\mathbf{I}_{N_t}E_{s}$, where $E_{s}$ denotes the symbol average energy, $\mathbb{E}[\cdot]$ denotes the expectation operation, $\mathbf{I}_{N_{t}}$ denotes identity matrix of size $N_{r}\times N_{t}$. Furthermore $\mathbf{H}\in \mathbb{C}^{N_{r}\times N_{t}}$ denotes the Rayleigh fading channel propagation matrix, each component is independent identically distributed (i.i.d) circularly symmetric complex Gaussian random variable with zero mean and unit variance. Finally, $\mathbf{n}\in \mathbb{C}^{N_{r}}$ is the additive white Gaussian noise (AWGN) vector with zero mean components and $\mathbb{E}[\mathbf{n}\mathbf{n}^{H}]=\mathbf{I}_{N_{r}}N_{0}$, where $N_{0}$ denotes the noise power spectrum density, and hence $\frac{E_{s}}{N_{0}}$ is the signal to noise ratio (SNR).

The task of a MIMO detector is to estimate the transmit symbol vector $\mathbf{s}$, based on the knowledge of receive symbol vector $\mathbf{y}$ and channel matrix $\mathbf{H}$.
\subsection{Complex Regression Model}
 Based on the discrete time model of the large MIMO uplink system of (\ref{discrete time MIMO system}), our regression model is set such that the training data samples at the detector are $(\mathbf{h}_{1}, y_{1})(\mathbf{h}_{2}, y_{2}), \ldots, (\mathbf{h}_{N_{r}}, y_{N_{r}})$, where $\mathbf{h}_{i}$ denotes $i$th row of matrix $\mathbf{H}$. This yields a regression task without offset $b$ in (\ref{equation1}) :  
 \begin{IEEEeqnarray}[\relax]{l}
 y_{i}=f_{true}(\mathbf{h}_{i})+n_{i}, i=1,2\ldots L\\
 \label{regression part1a}
 f_{true}(\mathbf{h}_{i})=\mathbf{h}_{i}\cdot\mathbf{s}, i=1,2,\ldots, L\\
 \label{regression part1b}
 \end{IEEEeqnarray}
 where $f_{true}()$ denotes the underlying true regression function, $n_{i}$ denotes discrete sample of additive noise.
In this regression problem, received symbols $y_{i}$ are the output data, $\mathbf{h}_{i}$ are the input data samples, transmitted symbol vector $\mathbf{s}$ contains the regression coefficients. We employ complex support vector regression (CSVR) without offset term $b$, in order to deal with the complex large MIMO systems. As shown in section \ref{Introduce epsilon SVR}, As shown in (\ref{original KKT condition1}) the first KKT condition is satisfied by finding the saddle point $\bar{\mathbf{w}}$ by taking paritial derivatives of $L(\mathbf{w},a_{i})$ with respect to $\mathbf{w}$. Mathematical results of Wirtinger's calculus in Reproducing Kernel Hilbert Space (RKHS) are employed to calculate partial derivatives of Lagrangian function of CSVR in complex manner\cite{bouboulis2011extension}. First we generalize our regression model by complex RKHS, 
Define a real RKHS $\mathcal{H}$, $<,>_{\mathcal{H}}$ denotes the corresponding inner product in $\mathcal{H}$. define complex RKHS $\mathbb{H}=\{f=f^{r}+\imath f^{i}, f^{r}, f^{i}\in \mathcal{H}\}$, $<,>_{\mathbb{H}}$ denotes the corresponding inner products in $\mathbb{H}$. Assume $\mathbf{x}$, $\mathbf{y}$, $\mathbf{z}$, $\mu$, $\chi$ $\in \mathbb{C}$, complex Hilbert space has the following properties 
\newtheorem{Lemma}{Lemma}
\newtheorem{Property}{Property}
\begin{Property}
 $<\mathbf{x},\mathbf{y}>_{\mathbb{H}}=(<{\mathbf{y},\mathbf{x}}>_{\mathbb{H}})^{*}$
\label{CHSProperty1}
\end{Property}

\begin{Property}
$<\mu\mathbf{x}+\chi\mathbf{y},\mathbf{z}>_{\mathbb{H}}=\mu<\mathbf{x},\mathbf{z}>_{\mathbb{H}}+\chi<\mathbf{y},\mathbf{z}>_{\mathbb{H}}$
\label{CHSProperty2}
\end{Property}

\begin{Property}
$<\mathbf{z}, \mu\mathbf{x}+\chi\mathbf{y}>_{\mathbb{H}}=\mu^{*}<\mathbf{z},\mathbf{x}>_{\mathbb{H}}+\chi^{*}<\mathbf{z}, \mathbf{y}>_{\mathbb{H}}$
\label{CHSProperty3}
\end{Property}

\begin{Lemma}
Assume $\mathbf{a}$, $\mathbf{b}$ $\in \mathbb{R}^{V}$, define a real RKHS $\mathcal{H}$ that satisfy $<\mathbf{a}, \mathbf{b}>_{\mathcal{H}}=\mathbf{a}^{T}\mathbf{b}$. For the corresponding complex RKHS $\mathbb{H}=\{f=f^{r}+\imath f^{i}, f^{r}, f^{i}\in \mathcal{H}\}$,
$<\mathbf{m},\mathbf{l}^{*}>_{\mathbb{H}}=\mathbf{m}\cdot \mathbf{l}$
\end{Lemma}

\begin{proof} 
$\mathbf{g}$,$\mathbf{h}$ $\in \mathbb{C}^{V}$, and $\mathbf{g}=\Re{(\mathbf{g})}+\imath\Im{(\mathbf{g})}$, $\mathbf{h}=\Re{(\mathbf{h})}+\imath\Im{(\mathbf{h})}$. From Property \ref{CHSProperty1} and Property \ref{CHSProperty3},  
\begin{IEEEeqnarray}[\relax]{l}
<\mathbf{g},\mathbf{h}>_{\mathbb{H}}=<\Re{(g)}+\imath\Im{(g)}, \Re{(h)}+\imath\Im{(h)}>_{\mathbb{H}}
\label{CHSProperty4 inter1}
\end{IEEEeqnarray}
According to Property \ref{CHSProperty2}, (\ref{CHSProperty4 inter1}) can be rewritten as 
\begin{equation}
<\mathbf{g},\mathbf{h}>_{\mathbb{H}}=<\Re{(\mathbf{g})}, \Re{(\mathbf{h})}+\imath\Im{(\mathbf{h})}>_{\mathbb{H}}+\imath <\Im{(\mathbf{g})}, \Re{(\mathbf{h})}+\imath\Im{(\mathbf{h})}>_{\mathbb{H}},
\label{CHSProperty4 inter2}
\end{equation}
According to Property \ref{CHSProperty3}
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
<\mathbf{g},\mathbf{h}>_{\mathbb{H}}=<\Re{(\mathbf{g})}, \Re{(\mathbf{h})}>_{\mathbb{H}}-\imath <\Re{(\mathbf{g})}, \Im{(\mathbf{h})}>_{\mathbb{H}}+\imath (<\Im{(\mathbf{g})}, \Re{(\mathbf{h})}>_{\mathbb{H}}-
\imath <\Im{(\mathbf{g})}, \\ \nonumber  \Im{(\mathbf{h})}>_{\mathbb{H}})
=<\Re{(\mathbf{g})},\Re{(\mathbf{h})}>_{\mathbb{H}}+<\Im{(\mathbf{g})},\Im{(\mathbf{h})}>_{\mathbb{H}}+\imath [<\Im{(\mathbf{g})},\Re{(\mathbf{h})}>_{\mathbb{H}}-<\Re{(\mathbf{g})},\\ \Im{(\mathbf{h})}>_{\mathbb{H}}]
\label{CHSProperty4 inter2}
\end{IEEEeqnarray}
Because $<\mathbf{f}_{1}, \mathbf{f}_{2}>_{\mathbb{H}}=<\mathbf{f}_{1}, \mathbf{f}_{2}>_{\mathcal{H}}=\mathbf{f}^{T}_{1}\cdot \mathbf{f}_{2}$ if $\mathbf{f}_{1}, \mathbf{f}_{2}\in \mathcal{H}$, thus (\ref{CHSProperty4 inter2}) can be rewritten as 
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
<\mathbf{g},\mathbf{h}>_{\mathbb{H}}=<\Re{(\mathbf{g})},\Re{(\mathbf{h})}>_{\mathcal{H}}+<\Im{(\mathbf{g})},\Im{(\mathbf{h})}>_{\mathcal{H}}+\imath [<\Im{(\mathbf{g})},\Re{(\mathbf{h})}>_{\mathcal{H}}-<\Re{(\mathbf{g})}, \Im{(\mathbf{h})}>_{\mathcal{H}}]=\\
\Re{(\mathbf{g})}^{T}\cdot \Re{(\mathbf{h})}+\Im{(\mathbf{g})}^{T}\cdot \Im{(\mathbf{h})}+\imath [\Im{(\mathbf{g})}^{T}\cdot  \Re{(\mathbf{h})}-\Re{(\mathbf{g})}^{T}\cdot \Im{(\mathbf{h})}]=\mathbf{g}\cdot \mathbf{h}^{*}
\label{CHSequation2}
\end{IEEEeqnarray}

Therefore $<\mathbf{m},\mathbf{l}>_{\mathbb{H}}=\mathbf{m}\cdot \mathbf{l}$.
\end{proof}
Therefore we have $<\mathbf{h}_{i}, \mathbf{s}^{*}>_{\mathbb{H}}=\mathbf{h}_{i}\cdot \mathbf{s}$, 
For sake of simplicity we define $\mathbf{s}^{*}=\mathbf{w}$. Based on the same principle in the real case presented in (\ref{Introduce epsilon SVR}), the general regularized risk function of large MIMO detection in complex RKHS $\mathbb{H}$ defined by $<\mathbf{m}, \mathbf{l}>_{\mathbb{H}}=\mathbf{m}^{T}\mathbf{l}$ can be formulated as 
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
\min_{\mathbf{w}, \xi^{r}_{i}, \hat{\xi}^{r}_{i}, \xi^{i}_{i}, \hat{\xi}^{i}_{i}} \quad \frac{1}{2}||\mathbf{w}||_{\mathbb{H}}^{2}+C\sum_{i=1}^{N_{r}}[(R(\xi^{r}_{i})+R(\hat{\xi}^{r}_{i})+R(\xi^{i}_{i})+R(\hat{\xi}^{i}_{i}))]\\
s.t. \left\{\begin{array}{ll}
\Re{(y_{i}-<\mathbf{h}_{i},\mathbf{w}>_{\mathbb{H}})}\leq \epsilon+\xi^{r}_{i}, i=1,2,\ldots, N_{r}\\
\Re{(<\mathbf{h}_{i},\mathbf{w}>_{\mathbb{H}}-y_{i})}\leq \epsilon+\hat{\xi}^{r}_{i}, i=1,2,\ldots, N_{r}\\
\Im{(y_{i}-<\mathbf{h}_{i},\mathbf{w}>_{\mathbb{H}})}\leq \epsilon+\xi^{i}_{i}, i=1,2,\ldots, N_{r}\\
\Im{(<\mathbf{h}_{i},\mathbf{w}>_{\mathbb{H}}-y_{i})}\leq \epsilon+\hat{\xi}^{i}_{i}, i=1,2,\ldots, N_{r}\\
\xi^{r}_{i}, \hat{\xi}^{r}_{i},\xi^{i}_{i},\hat{\xi}^{i}_{i}\geq 0, i=1,2,\ldots, N_{r}\\
\end{array}\right.
\label{complex objective function1}
\end{IEEEeqnarray}
 Where $\Re{(\cdot)}$ and $\Im{(\cdot)}$ denote real and imaginary part of a complex variable and inequality restrictions are set to real and imaginary part of the regression function separately. Let $\mathbf{K}=\mathbf{H}\mathbf{H}^{H}$ denote the kernel function, $\mathbf{K}=\Re{(\mathbf{K})}+\imath\Im{(\mathbf{K})}$, where $\Re{(\mathbf{K})}$ and $\Im{(\mathbf{K})}$ denote matrices of corresponding real and imaginary parts. Similar to the Lagrange duality rational in section \ref{section lagrange duality}, the Lagrange function associated with (\ref{complex objective function1}) is 
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
L=
\frac{1}{2}||\mathbf{w}||_{\mathbb{H}}^{2}+C\sum_{i=1}^{N_{r}}[(R(\xi^{r}_{i})+R(\hat{\xi}^{r}_{i})+R(\xi^{i}_{i})+R(\hat{\xi}^{i}_{i}))]- \sum_{i=1}^{N_{r}}(\eta_{i}\xi^{r}_{i}+\hat{\eta}_{i}\hat{\xi}^{r}_{i}+\tau_{i}\xi^{i}_{i}\\\nonumber +\hat{\tau}_{i}\hat{\xi}^{i}_{i})
+\sum_{i=1}^{N_{r}}\alpha_{i}(\Re{(y_{i})}-\Re{(<\mathbf{h}_{i},\mathbf{w}>_{\mathbb{H}})}-\epsilon-\xi^{r}_{i})
+\sum_{i=1}^{N_{r}}\hat{\alpha}_{i}(\Re{(<\mathbf{h}_{i},\mathbf{w}>_{\mathbb{H}})}-\Re{(y_{i})}-\epsilon-\hat{\xi}^{r}_{i})\\
\nonumber
+\sum_{i=1}^{N_{r}}\beta_{i}(\Im{(y_{i})}-\Im{(<\mathbf{h}_{k},\mathbf{w}>_{\mathbb{H}})}-\epsilon-\xi^{i}_{i})+\sum_{i=1}^{N_{r}}\hat{\beta}_{i}(\Im{(<\mathbf{h}_{k},\mathbf{w}>_{\mathbb{H}})}-\Im{(y_{i})}-\epsilon-\hat{\xi}^{i}_{i})\\
s.t. \left\{\begin{array}{ll}
\eta_{i}, \hat{\eta}_{i}, \tau_{i}, \hat{\tau}_{i}, \alpha_{i}, \hat{\alpha}_{i}, \beta_{i}, \hat{\beta}_{i}\geq 0, i=1,2,\ldots, N_{r}\\
\xi^{r}_{i}, \hat{\xi}^{r}_{i}, \xi^{i}_{i}, \hat{\xi}^{i}_{i} \geq  0, i=1,2,\ldots, N_{r}
\end{array}\right. 
\label{complex lagrange duality1}
\end{IEEEeqnarray}
where $\eta_{i}, \hat{\eta}_{i}, \tau_{i}, \hat{\tau}_{i}, \alpha_{i}, \hat{\alpha}_{i}, \beta_{i}, \hat{\beta}_{i}$ are Lagrange multipliers. In order to calculate the partial derivations of $L$ with respect to $\mathbf{w}$ which is defined in complex domain, the Wirtinger's calculus is exploited\cite{bouboulis2011extension} , Therefore we have   
\begin{IEEEeqnarray}[\relax]{l}
\left\{\begin{array}{ll}
\frac{\partial L}{\partial \mathbf{w}^{*}}=\frac{1}{2}\mathbf{w}-\frac{1}{2}\sum_{i=1}^{N_{r}}\alpha_{i}\mathbf{h}_{i}+\frac{1}{2}\sum_{i=1}^{N_{r}}\hat{\alpha}_{i}\mathbf{h}_{i}+\frac{\imath}{2}(\sum_{i=1}^{N_{r}}\beta_{i}\mathbf{h}_{i}-\sum_{i=1}^{N_{r}}\hat{\beta}_{i}\mathbf{h}_{i})=0\\
\Rightarrow \mathbf{w}=\sum_{i=1}^{N_{r}}(\alpha_{i}-\hat{\alpha}_{i})\mathbf{h}_{i}-\imath\sum_{i=1}^{N_{r}}(\beta_{i}-\hat{\beta}_{i})\mathbf{h}_{i}\\
\frac{\partial L}{\partial \xi^{r}_{i}}=CR^{'}(\xi^{r}_{i})-\eta_{i}-\alpha_{i}=0\Rightarrow \eta_{i}=CR^{'}(\xi^{r}_{i})-\alpha_{i}\\
\frac{\partial L}{\partial \hat{\xi}^{r}_{i}}=CR^{'}(\hat{\xi}^{r}_{i})-\hat{\eta}_{i}-\hat{\alpha}_{i}=0\Rightarrow \hat{\eta}_{i}=CR^{'}(\hat{\xi}^{r}_{i})-\hat{\alpha}_{i}\\
\frac{\partial L}{\partial \xi^{i}_{i}}=CR^{'}(\xi^{i}_{i})-\tau_{i}-\beta_{i}=0\Rightarrow \tau_{i}=CR^{'}(\xi^{i}_{i})-\beta_{i}\\
\frac{\partial L}{\partial \hat{\xi}^{i}_{i}}=CR^{'}(\hat{\xi}^{i}_{i})-\hat{\tau}_{i}-\hat{\beta}_{i}=0\Rightarrow \hat{\tau}_{i}=CR^{'}(\hat{\xi}^{i}_{i})-\hat{\beta}_{i}\\

\end{array}\right.
\label{complex lagrange duality2}
\end{IEEEeqnarray}
Based on (\ref{complex lagrange duality2}), we have 
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
<\mathbf{h}_{i}, \mathbf{w}>_{\mathbb{H}}=\sum_{j=1}^{N_{r}}(\alpha_{j}-\hat{\alpha}_{j})<\mathbf{h}_{i},\mathbf{h}_{j}>_{\mathbb{H}}+i\sum_{j=1}^{N_{r}}(\beta_{j}-\hat{\beta}_{j})<\mathbf{h}_{i},\mathbf{h}_{j}>_{\mathbb{H}}\\
\nonumber
=\sum_{j=1}^{N_{r}}(\alpha_{j}-\hat{\alpha}_{j})\mathbf{h}_{i}\mathbf{h}^{H}_{j}+i\sum_{j=1}^{N_{r}}(\beta_{j}-\hat{\beta}_{j})\mathbf{h}_{i}\mathbf{h}^{H}_{j}\\\nonumber
=\sum_{j=1}^{N_{r}}(\alpha_{j}-\hat{\alpha}_{j})\Re{(\mathbf{K})}_{ij}-\sum_{j=1}^{N_{r}}(\beta_{j}-\hat{\beta}_{j})\Im{(\mathbf{K})}_{ij}+i(\sum_{j=1}^{N_{r}}(\alpha_{j}-\hat{\alpha}_{j})\Im{(\mathbf{K})}_{ij}+\\ \sum_{j=1}^{N_{r}}(\beta_{j}-\hat{\beta}_{j})\Re{(\mathbf{K})}_{ij}),
\label{complex duality regression function}
\end{IEEEeqnarray}

\begin{IEEEeqnarray}[\relax]{l}
\nonumber
||\mathbf{w}||^{2}_{\mathbb{H}}=\sum_{i,j=1}^{N_{r}}(\alpha_{i}-\hat{\alpha}_{i})(\alpha_{i}-\hat{\alpha}_{i})\mathbf{h}_{i}\mathbf{h}^{H}_{j}+\sum_{i,j=1}^{N_{r}}(\beta_{i}-\hat{\beta}_{i})(\beta_{i}-\hat{\beta}_{i})\mathbf{h}_{i}\mathbf{h}^{H}_{j}\\
+i(\sum_{i,j=1}^{N_{r}}(\alpha_{i}-\hat{\alpha}_{i})(\beta_{j}-\hat{\beta}_{j})\mathbf{h}_{i}\mathbf{h}^{H}_{j}-\sum_{i,j=1}^{N_{r}}(\alpha_{i}-\hat{\alpha}_{i})(\beta_{j}-\hat{\beta}_{j})\mathbf{h}_{j}\mathbf{h}^{H}_{i})
\label{complex duality regularization1}
\end{IEEEeqnarray}
Because $\mathbf{K}$ is Hermitian, thus $\mathbf{K}_{ij}=\mathbf{K}^{*}_{ji}$. If we have $r_{i}$, $r_{j}$ $\in \mathbb{R}$, then 
\begin{equation}
\sum_{i,j}^{L}r_{i}r_{j}\Im{(\mathbf{K})}_{ij}=-\sum_{i,j}^{L}r_{i}r_{j}\Im{(\mathbf{K})}_{ji}=-\sum_{i,j}^{L}r_{i}r_{j}\Im{(\mathbf{K})}_{ij},
\end{equation}
Therefore
\begin{equation}
\sum_{i,j}^{l}r_{i}r_{j}\Im{(\mathbf{K})}_{ij}=0,
\label{conjugate1}
\end{equation}
Based on (\ref{conjugate1}), (\ref{complex duality regularization1}) can be changed to
\begin{equation}
||\mathbf{w}||^{2}_{\mathbb{H}}=\sum_{i,j=1}^{N_{r}}(\alpha_{i}-\hat{\alpha}_{i})(\alpha_{j}-\hat{\alpha}_{j})\Re{(\mathbf{K})}_{ij}+\sum_{i,j=1}^{N_{r}}(\beta_{i}-\hat{\beta}_{i})(\beta_{j}-\hat{\beta}_{j})\Re{(\mathbf{K})}_{ij}-2\sum_{i,j=1}^{N_{r}}(\alpha_{i}-\hat{\alpha}_{i})(\beta_{j}-\hat{\beta}_{j})\Im{(\mathbf{K})}_{ij}.
\label{complex duality regularization2}
\end{equation}
Apply (\ref{complex lagrange duality2}), (\ref{complex duality regression function}), (\ref{conjugate1}) and (\ref{complex duality regularization2}) to (\ref{complex lagrange duality1}), the final form of Lagrange duality can be obtained
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
L=\frac{1}{2}[\sum_{i,j=1}^{N_{r}}(\alpha_{i}-\hat{\alpha}_{i})(\alpha_{j}-\hat{\alpha}_{j})\Re{(\mathbf{K})}_{ij}+\sum_{i,j=1}^{N_{r}}(\beta_{i}-\hat{\beta}_{i})(\beta_{j}-\hat{\beta}_{j})\Re{(\mathbf{K})}_{ij}-2\sum_{i,j=1}^{N_{r}}(\alpha_{i}-\hat{\alpha}_{i})(\beta_{j}-\hat{\beta}_{j})\Im{(\mathbf{K})}_{ij}]+\\
\nonumber
C\sum_{i=1}^{N_{r}}[R(\xi^{r}_{i})+R(\hat{\xi}^{r}_{i})+R(\xi^{i}_{i})+R(\hat{\xi}^{i}_{i})]- \sum_{i=1}^{N_{r}}((CR^{'}(\xi^{r}_{i})-\alpha_{i})\xi^{r}_{i}+(CR^{'}(\hat{\xi}_{i}^{r})-\hat{\alpha}_{i})\hat{\xi}^{r}_{i}+(CR^{'}-\beta_{i})\xi^{i}_{i}\\\nonumber +(CR^{'}(\hat{\xi}_{i})-\hat{\beta}_{i})\hat{\xi}^{i}_{i})+\sum_{i=1}^{N_{r}}\alpha_{i}(\Re{(y_{i})}-(\sum_{j=1}^{N_{r}}(\alpha_{j}-\hat{\alpha}_{j})\Re{(\mathbf{K})}_{ij}-\sum_{j=1}^{N_{r}}(\beta_{j}-\hat{\beta}_{j})\Im{(\mathbf{K})}_{ij})-\epsilon-\xi^{r}_{i})
+\sum_{i=1}^{N_{r}}\hat{\alpha}_{i}\\ \nonumber
((\sum_{j=1}^{N_{r}}(\alpha_{j}-\hat{\alpha}_{j})\Re{(\mathbf{K})}_{ij}-\sum_{j=1}^{N_{r}}(\beta_{j}-\hat{\beta}_{j})\Im{(\mathbf{K})}_{ij})-\Re{(y_{i})}-\epsilon-\hat{\xi}^{r}_{i})
+\sum_{i=1}^{N_{r}}\beta_{i}(\Im{(y_{i})}-(\sum_{j=1}^{N_{r}}(\alpha_{j}-\hat{\alpha}_{j})\\ \nonumber
\Im{(\mathbf{K})}_{ij}+\sum_{j=1}^{N_{r}}(\beta_{j}-\hat{\beta}_{j})\Re{(\mathbf{K})}_{ij})-\epsilon-\xi^{i}_{i})+\sum_{i=1}^{N_{r}}\hat{\beta}_{i}((\sum_{j=1}^{N_{r}}(\alpha_{j}-\hat{\alpha}_{j})\Im{(\mathbf{K})}_{ij}+\sum_{j=1}^{N_{r}}(\beta_{j}-\hat{\beta}_{j})\Re{(\mathbf{K})}_{ij})-\Im{(y_{i})}-\\
\epsilon-\hat{\xi}^{i}_{i})
\label{complex lagrange duality inter1}
\end{IEEEeqnarray}

\begin{IEEEeqnarray}[\relax]{l}
\nonumber
\theta=-\frac{1}{2}[\sum_{i,j=1}^{N_{r}}(\alpha_{i}-\hat{\alpha}_{i})(\alpha_{j}-\hat{\alpha}_{j})\Re{(\mathbf{K})}_{ij}+\sum_{i,j=1}^{N_{r}}(\beta_{i}-\hat{\beta}_{i})(\beta_{j}-\hat{\beta}_{j})\Re{(\mathbf{K})}_{ij}]+C\sum_{i=1}^{N_{r}}[R(\xi^{r}_{i})-\xi_{i}^{r}R^{'}(\xi^{r}_{i})\\\nonumber
+R(\hat{\xi}^{r}_{i})-\hat{\xi}^{r}_{i}R^{'}(\hat{\xi}^{r}_{i})+R(\xi^{i}_{i})-\xi_{i}R^{'}(\xi_{i}^{i})+R(\hat{\xi}^{i}_{i})-\hat{\xi}^{i}_{i}R^{'}(\hat{\xi}^{i}_{i})]+\sum_{i=1}^{N_{r}}[\Re{(y_{i})}(\alpha_{i}-\hat{\alpha}_{i})+\Im{(y_{i})}(\beta_{i}-\hat{\beta}_{i})]-\\
\epsilon\sum_{i=1}^{N_{r}}(\alpha_{i}+\hat{\alpha}_{i}+\beta_{i}+\hat{\beta}_{i})
\label{complex lagrange duality inter2}
\end{IEEEeqnarray}
define $\tilde{R}(u)=R(u)-uR^{'}(u)$. Similar to (\ref{dual optimization function2}), The dual optimization problem of complex MIMO system is 
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
\max_{\alpha_{i}, \hat{\alpha}_{i}, \beta_{i}, \hat{\beta}_{i}}\quad \theta=-\frac{1}{2}[\sum_{i,j}^{N_{r}}(\alpha_{i}-\hat{\alpha}_{i})(\alpha_{j}-\hat{\alpha}_{j})\mathbf{K}^{r}_{ij}+\sum_{i,j}^{N_{r}}(\beta_{i}-\hat{\beta}_{i})(\beta_{j}-\hat{\beta}_{j})\mathbf{K}^{r}_{ij}]\\
\nonumber
-\sum_{i}^{N_{r}}(\alpha_{i}+\hat{\alpha}_{i}+\beta+\hat{\beta}_{i})\epsilon+[\sum_{i=1}^{N_{r}}(\alpha_{i}-\hat{\alpha}_{i})Re(y_{i})+\sum_{i=1}^{N_{r}}(\beta_{i}-\hat{\beta}_{i})Im(y_{i})]\\
\nonumber
+C\sum_{i}^{N_{r}}(\tilde{R}(\xi^{r}_{i})+\tilde{R}(\hat{\xi}^{r}_{i})+\tilde{R}(\xi^{i}_{i})+\tilde{R}(\hat{\xi}^{i}_{i}))\\ 
\left\{\begin{array}{ll}
0\leq \alpha_{i}(\hat{\alpha}_{i})\leq C\tilde{R}(\xi_{i}^{r})(\tilde{R}(\hat{\xi}_{i}^{r})), i=1,2,\ldots, L\\
0\leq \beta_{i}(\hat{\beta}_{i})\leq C\tilde{R}(\xi_{i}^{i})(\tilde{R}(\hat{\xi}_{i}^{i})), i=1,2,\ldots, L\\
\xi_{i}^{r}(\hat{\xi}_{i}^{r})\geq 0, i=1,2,\ldots, L\\
\xi_{i}^{i}(\hat{\xi}_{i}^{i})\geq 0, i=1,2,\ldots, L\\
\end{array}\right.
\label{final complex lagrange duality}
\end{IEEEeqnarray}
Notice in (\ref{final complex lagrange duality}), there is no correlation term between $\alpha_{i}, \hat{\alpha}_{i}$ and $\beta_{i}, \hat{\beta}_{i}$, therefore, (\ref{final complex lagrange duality}) can be divided into two independent regression tasks,

\begin{IEEEeqnarray}[\relax]{l}
\nonumber
\max_{\alpha_{i}, \hat{\alpha}_{i}} \quad \theta^{r}= -\frac{1}{2}\sum_{i,j}^{N_{r}}(\alpha_{i}-\hat{\alpha}_{i})(\alpha_{j}-\hat{\alpha}_{j})\Re{(\mathbf{K})}_{ij}-\sum_{i=1}^{N_{r}}(\alpha_{i}+\hat{\alpha}_{i})\epsilon+\sum_{i=1}^{N_{r}}(\alpha_{i}-\hat{\alpha}_{i})Re(y_{i})+C\sum_{i=1}^{N_{r}}(\tilde{R}(\xi^{r}_{i})\\
\nonumber
+\tilde{R}(\hat{\xi}^{r}_{i}))\\
\left\{\begin{array}{ll}
0\leq \alpha_{i}(\hat{\alpha}_{i})\leq C\tilde{R}(\xi_{i}^{r})(\tilde{R}(\hat{\xi}_{i}^{r})), i=1,2,\ldots, L\\
\xi_{i}^{r}(\hat{\xi}_{i}^{r})\geq 0, i=1,2,\ldots, L\\
\end{array}\right.
\label{complex duality real part}
\end{IEEEeqnarray}

\begin{IEEEeqnarray}[\relax]{l}
\nonumber
\max_{\beta_{i}, \hat{\beta}_{i}} \quad \theta^{i}= -\frac{1}{2}\sum_{i,j}^{N_{r}}(\beta_{i}-\hat{\beta}_{i})(\beta_{j}-\hat{\beta}_{j})\Re{(\mathbf{K})}_{ij}-\sum_{i=1}^{N_{r}}(\beta_{i}+\hat{\beta}_{i})\epsilon+\sum_{i=1}^{N_{r}}(\beta_{i}-\hat{\beta}_{i})Im(y_{i})+C\sum_{i=1}^{N_{r}}(\tilde{R}(\xi^{i}_{i})\\
\nonumber
+\tilde{R}(\hat{\xi}^{i}_{i}))\\
\left\{\begin{array}{ll}
0\leq \beta_{i}(\hat{\beta}_{i})\leq C\tilde{R}(\xi_{i}^{i})(\tilde{R}(\hat{\xi}_{i}^{i})), i=1,2,\ldots, L\\
\xi_{i}^{i}(\hat{\xi}_{i}^{i})\geq 0, i=1,2,\ldots, L\\
\end{array}\right.
\label{complex duality imaginary part}
\end{IEEEeqnarray}
Let $\mathbf{a}=[\alpha_{1}, \alpha_{2},\ldots, \alpha_{N_{r}}]^{T}$, $\mathbf{\hat{a}}=[\hat{\alpha}_{1}, \hat{\alpha}_{2},\ldots, \hat{\alpha}_{N_{r}}]^{T}$, $\mathbf{b}=[\beta_{1}, \beta_{2},\ldots, \beta_{N_{r}}]^{T}$,$\mathbf{\hat{b}}=[\hat{\beta}_{1}, \hat{\beta}_{2},\ldots, \hat{\beta}_{N_{r}}]^{T}$, $\Re{(\mathbf{y})}=[\Re{(y_{1})}, \Re{(y_{2})},\ldots, \Re{(y_{N_{r}})}]^{T}$, $\Im{(\mathbf{y})}=[\Im{(y_{1})}, \Im{(y_{2})}, \ldots, \Im{(y_{N_{r}})}]^{T}$, $\mathbf{R}^{r}=[\tilde{R}(\xi^{r}_{1})+\tilde{R}(\hat{\xi}^{r}_{1}), \tilde{R}(\xi^{r}_{2})+\tilde{R}(\hat{\xi}^{r}_{2}), \ldots, \tilde{R}(\xi^{r}_{N_{r}})+\tilde{R}(\hat{\xi}^{r}_{N_{r}})]^{T}$, $\mathbf{R}^{i}=[\tilde{R}(\xi^{i}_{1})+\tilde{R}(\hat{\xi}^{i}_{1}), \tilde{R}(\xi^{i}_{2})+\tilde{R}(\hat{\xi}^{i}_{2}), \ldots, \tilde{R}(\xi^{i}_{N_{r}})+\tilde{R}(\hat{\xi}^{i}_{N_{r}})]^{T}$, $\mathbf{e}=[1,1,\ldots, 1]^{T}\in \mathbb{R}^{N_{r}}$. The alternate form can be written as 
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
\max_{\mathbf{a},\mathbf{\hat{a}}} \quad \theta^{r}=-\frac{1}{2}(\mathbf{a}-\mathbf{\hat{a}})^{T}\Re{(\mathbf{K})}(\mathbf{a}-\mathbf{\hat{a}})+\Re{(\mathbf{y})}^{T}(\mathbf{a}-\mathbf{\hat{a}})-\epsilon(\mathbf{e}^{T}(a+\mathbf{\hat{a}}))+C(\mathbf{e}^{T}\mathbf{R}^{r})\\
\left\{\begin{array}{ll}
0\leq \alpha_{i}(\hat{\alpha}_{i})\leq C\tilde{R}(\xi_{i}^{r})(\tilde{R}(\hat{\xi}_{i}^{r})), i=1,2,\ldots, N_{r}\\
\xi_{i}^{r}(\hat{\xi}_{i}^{r})\geq 0, i=1,2,\ldots, N_{r}\\
\end{array}\right.
\label{complex duality real part vectorizes}
\end{IEEEeqnarray}

\begin{IEEEeqnarray}[\relax]{l}
\nonumber
\max_{\mathbf{b}, \mathbf{\hat{b}}} \quad \theta^{i}=-\frac{1}{2}(\mathbf{b}-\mathbf{\hat{b}})^{T}\Re{(\mathbf{K})}(\mathbf{b}-\mathbf{\hat{b}})+\Im{(\mathbf{y})}^{T}(\mathbf{b}-\mathbf{\hat{b}})-\epsilon(\mathbf{e}^{T}(\mathbf{b}+\mathbf{\hat{b}}))+C(\mathbf{e}^{T}\mathbf{R}^{i})\\
\left\{\begin{array}{ll}
0\leq \beta_{i}(\hat{\beta}_{i})\leq C\tilde{R}(\xi_{i}^{i})(\tilde{R}(\hat{\xi}_{i}^{i})), i=1,2,\ldots, N_{r}\\
\xi_{i}^{i}(\hat{\xi}_{i}^{i})\geq 0, i=1,2,\ldots, N_{r}\\
\end{array}\right.
\label{complex duality imaginary part vectorized}
\end{IEEEeqnarray}

Observe that solving (\ref{complex duality real part vectorizes}) and (\ref{complex duality imaginary part vectorized}) are equivalent to solving two independent real Support vector regression task (dual channel), where only the real part of the kernel matrix is required for each channel. This can provide computational advantages for solving the dual optimization problems.    
\section{Work Set Selection and Solver}\label{WSS}
In (\ref{final complex lagrange duality}) we have a quadratic optimization problem, The traditional optimization algorithms such as Newton, Quasi Newton can not be directly applied to this problem, because the sparseness of kernel matrix $\mathbf{K}$ can not be guaranteed, so that a prohibitive storage may be required when dealing with a large data set.

Decomposition methods are a set of efficient algorithms that can help to ease this difficulty. Decomposition methods work iteratively, and based on choosing a subset of variables $S$ (named work set) to optimize in each iteration step, while keeping the rest variables $N$ fixed. Sequential Minimal Optimization (SMO) is an extreme case of decomposition method, where the work set size is 2, and an analytic quadratic programming (QP) step instead of numerical QP step can be performed in each iteration.

Because (\ref{complex duality real part}) and (\ref{complex duality imaginary part}) have the same structure, in this section we discuss the real part only. By dividing the variables into work set $S$ and fixed set $N$, we can divide vector $\mathbf{a}$ into two sub vectors $[\mathbf{a}_{S}, \mathbf{a}_{N}]$.  thereafter $\mathbf{a}_{S}$ denotes a vector that consists of the components of $\mathbf{a}$ that belongs to set $S$, the same modifications are applied to $\mathbf{\hat{a}}$ and $\mathbf{y}$. Thus (\ref{complex duality real part vectorizes}) can be rewritten as:
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
\max_{\mathbf{a}_{S}, \hat{\mathbf{a}}_{S}, \mathbf{a}_{N}, \hat{\mathbf{a}}_{N}} \theta= -\frac{1}{2}[(\mathbf{a}_{S}-\hat{\mathbf{a}}_{S})^{T}, (\mathbf{a}_{N}-\hat{\mathbf{a}}_{N})^{T}]\left[\begin{IEEEeqnarraybox*}[\mysmallarraydecl][c]{,c/c,}
\Re{(\mathbf{K})}_{SS}& \Re{(\mathbf{K})}_{SN}\\
\Re{(\mathbf{K})}_{NS}& \Re{(\mathbf{K})}_{NN}
\end{IEEEeqnarraybox*}\right]\left[\begin{IEEEeqnarraybox*}[\mysmallarraydecl][c]{,c/c,}
(\mathbf{a}_{S}-\hat{\mathbf{a}}_{S})\\
(\mathbf{a}_{N}-\hat{\mathbf{a}}_{N})
\end{IEEEeqnarraybox*}\right]+[\Re{(\mathbf{y}_{S})}^{T}, \Re{(\mathbf{y}_{N})}^{T}]\left[\begin{IEEEeqnarraybox*}[\mysmallarraydecl][c]{,c/c,}
(\mathbf{a}_{S}-\hat{\mathbf{a}}_{S})\\
(\mathbf{a}_{N}-\hat{\mathbf{a}}_{N})
\end{IEEEeqnarraybox*}\right]-\\\nonumber\epsilon[\mathbf{e}^{T}_{S}, \mathbf{e}^{T}_{N}]\left[\begin{IEEEeqnarraybox*}[\mysmallarraydecl][c]{,c/c,}
(\mathbf{a}_{S}+\hat{\mathbf{a}}_{S})\\
(\mathbf{a}_{N}+\hat{\mathbf{a}}_{N})
\end{IEEEeqnarraybox*}\right]+C(\mathbf{e}^{T}, \mathbf{R}^{r})
\\
\left\{\begin{array}{ll}
0\leq [\mathbf{a}]_{i}([\hat{\mathbf{a}}]_{i})\leq C\tilde{R}(\xi_{i}^{r})(\tilde{R}(\hat{\xi}_{i}^{r})), i=1,2,\ldots, N_{r}\\
\xi_{i}^{r}(\hat{\xi}_{i}^{r})\geq 0, i=1,2,\ldots, N_{r}\\
\end{array}\right.
\label{real part decomposition inter1}
\end{IEEEeqnarray}

Where $\left[\begin{IEEEeqnarraybox*}[\mysmallarraydecl][c]{,c/c,}
\Re{(\mathbf{K})}_{SS}& \Re{(\mathbf{K})}_{SN}\\
\Re{(\mathbf{K})}_{NS}& \Re{(\mathbf{K})}_{NN}
\end{IEEEeqnarraybox*}\right]$ is a permutation of $\Re{(\mathbf{K})}$, where $\Re{(\mathbf{K})_{SN}}$ denotes the sub matrix that consist of rows corresponding to set $S$ and columns corresponding to $N$, Notice that $\Re{(\mathbf{K})}_{SN}=\Re{(\mathbf{K})}_{NS}$, therefore, (\ref{real part decomposition inter1}) can be rewritten as 
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
\max_{\mathbf{a}, \mathbf{\hat{a}}}\quad \theta^{r}=-\frac{1}{2}[(\mathbf{a}_{S}-\mathbf{\hat{a}}_{S})^{T}\Re{(\mathbf{K})}_{SS}(\mathbf{a}_{S}-\mathbf{\hat{a}}_{S})+2(\mathbf{a}_{N}-\mathbf{\hat{a}}_{N})^{T}\Re{(\mathbf{K})}_{NS}(\mathbf{a}_{S}-\mathbf{\hat{a}}_{S})]
+\Re{(\mathbf{y}_{S}^{T})}(\mathbf{a}_{S}-\mathbf{\hat{a}}_{S})-\\
\nonumber
\epsilon(\mathbf{e}^{T}(\mathbf{a}_{S}+\hat{\mathbf{a}}_{S}))
-\frac{1}{2}(\mathbf{a}_{N}-\hat{\mathbf{a}}_{N})^{T}\Re{(\mathbf{K})}_{NN}(\mathbf{a}_{N}-\mathbf{\hat{a}}_{N})+\Re{(\mathbf{y}_{N}^{T})}(\mathbf{a}_{N}-\mathbf{\hat{a}}_{N})-\epsilon(\mathbf{e}^{T}(\mathbf{a}_{N}+\mathbf{\hat{a}}_{N}))\\\nonumber+C(\mathbf{e}^{T}(\mathbf{R}^{r}),
\\
\left\{\begin{array}{ll}
0\leq [\mathbf{a}]_{i}([\hat{\mathbf{a}}]_{i})\leq C\tilde{R}(\xi_{i}^{r})(\tilde{R}(\hat{\xi}_{i}^{r})), i=1,2,\ldots, N_{r}\\
\xi_{i}^{r}(\hat{\xi}_{i}^{r})\geq 0, i=1,2,\ldots, N_{r}\\
\end{array}\right.
\label{real part decomposition}
\end{IEEEeqnarray}
In each iteration, in (\ref{real part decomposition}), $\mathbf{a}_{N}$ is fixed and only the sub  QP problem that related to $\mathbf{a}_{S}$ is solved i.e 
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
\max_{\mathbf{a}_{S}, \mathbf{\hat{a}}_{S}}\quad \theta_{S}^{r}=-\frac{1}{2}[(\mathbf{a}_{S}-\mathbf{\hat{a}}_{S})^{T}\Re{(\mathbf{K})}_{SS}(\mathbf{a}_{S}-\mathbf{\hat{a}}_{S})]+[Re(\mathbf{y}_{S}^{T})-(\mathbf{a}_{N}-\mathbf{\hat{a}}_{N})^{T}\mathbf{K}_{NS}^{r}
](\mathbf{a}_{S}-\mathbf{\hat{a}}_{S})-\\\nonumber\epsilon<\mathbf{e}^{T}, (\mathbf{a}_{S}+\mathbf{\hat{a}}_{S})>,
\\
\left\{\begin{array}{ll}
0\leq [\mathbf{a}_{S}]_{i}([\hat{\mathbf{a}}_{S}]_{i})\leq C\tilde{R}(\xi_{i}^{r})(\tilde{R}(\hat{\xi}_{i}^{r})), i\in S\\
\xi_{i}^{r}(\hat{\xi}_{i}^{r})\geq 0, i\in S\\
\end{array}\right.
\label{subset optimization}
\end{IEEEeqnarray}
A proper work set selection strategy is required so that speed and performance requirement can be guaranteed. One idea is to perform update to dual variables that violate the complementary KKT conditions, which, based on (\ref{original KKT condition3}) in \ref{section lagrange duality}, can be formulated as 
\begin{IEEEeqnarray}[\relax]{l}
\left\{\begin{array}{ll}
(C\tilde{R}(\xi_{i}^{r})-\alpha_{i})\xi_{i}^{r}=0, i=1,2,\ldots, L\\
(C\tilde{R}(\hat{\xi}_{i}^{r})-\hat{\alpha}_{i})\hat{\xi}_{i}^{r}=0, i=1,2,\ldots, L\\
\alpha(\epsilon+\xi_{i}^{r}-\Re{(\mathbf{y}_{i})}+<\mathbf{h}_{i},\mathbf{w}>_{\mathbb{H}})=0, i=1,2,\ldots, L\\
\hat{\alpha}_{i}(\epsilon+\hat{\xi}_{i}^{r}-<\mathbf{h}_{i},\mathbf{w}>_{\mathbb{H}}+\Re{(\mathbf{y}_{i})})=0, i=1,2,\ldots, L\\
\end{array}\right.
\label{complementary KKT condition1}
\end{IEEEeqnarray}
According to Osuna's theorem\cite{osuna1997improved}, the final convergence can be guaranteed. In SMO algorithm, heuristic methods are used to find a work set of size two in order to accelerate the decomposition process\cite{platt1999fast}. The heuristic method first searches among the non-bound variables (that is $0<\alpha_{i}<C\tilde{R}(\xi^{r}_{i})$ and $0<\hat{\alpha}_{i}<C\tilde{R}(\hat{\xi}_{i})$), which are more likely to violate the complementary KKT conditions, Then searching the whole dual variable set. The second dual variable that can maximize optimization step size is chosen. An approximate step size is used as evaluator for sake of reducing computational cost.

Another idea for work set selection is to choose the dual variables whose update can provide the maximum improvements to dual objective function. That is
\begin{equation}
\bigtriangledown \theta^{r}_{S}=\theta^{r}_{S}((\mathbf{a}_{S}+\Delta_{S}), (\hat{\mathbf{a}}_{S}+\hat{\Delta}_{S}))-\theta^{r}_{S}(\mathbf{a}_{S}, \hat{\mathbf{a}}_{S}),
\label{maximum gain work set selection}
\end{equation} 
where $\Delta_{S}=\mathbf{a}_{S}^{new}-\mathbf{a}_{S}$, $\mathbf{a}_{S}$ is the update of $\mathbf{a}^{new}_{S}$. $\bigtriangledown \theta_{S}^{r}$ in (\ref{maximum gain work set selection}),  based on (\ref{subset optimization}), can be written as 
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
\bigtriangledown \theta_{S}^{r}=-\frac{1}{2}[(\Delta_{S}-\hat{\Delta}_{S})^{T}\Re{(\mathbf{K})}_{SS}(\mathfrak{n}_{S}-\hat{\Delta}_{S})+2(\mathbf{a}_{S}-\hat{\mathbf{a}}_{S})^{T}\Re{(\mathbf{K})}_{SS}(\Delta_{S}-\hat{\Delta})_{S}]+[\Re{(\mathbf{y}^{T}_{S})}-(\mathbf{a}_{N}-\hat{\mathbf{a}}_{N})^{T}\Re{(\mathbf{K})}_{NS}](\Delta_{S}-\hat{\Delta}_{S})\\
-\epsilon\mathbf{e}^{T}_{S}(\Delta_{S}+\hat{\Delta}_{S})=-\frac{1}{2}\Delta_{S}-\hat{\Delta}_{S})^{T}\Re{(\mathbf{K})}_{SS}(\Delta_{S}-\hat{\Delta}_{S})+[\Re{(\mathbf{y}^{T}_{S})}-(\mathbf{a}-\hat{\mathbf{a}})^{T}\Re{(\mathbf{K})}_{S}](\Delta_{S}-\hat{\Delta}_{S})\\
-\epsilon\mathbf{e}^{T}_{S}(\Delta_{S}+\hat{\Delta}_{S}),
\label{gain work set selection1}
\end{IEEEeqnarray} 
Notice in (\ref{gain work set selection1}), we use 
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
(\mathbf{a}_{S}-\hat{\mathbf{a}}_{S})^{T}\Re{(\mathbf{K})}_{SS}+(\mathbf{a}_{N}-\hat{\mathbf{a}}_{N})^{T}\Re{(\mathbf{K})}_{NS}=[(\mathbf{a}_{S}-\hat{\mathbf{a}}_{S})^{T}, (\mathbf{a}_{N}-\hat{\mathbf{a}}_{N})^{T}]\left[\begin{IEEEeqnarraybox*}[\mysmallarraydecl][c]{,c/c,}
\Re{(\mathbf{K})}_{SS}\\
\Re{(\mathbf{K})}_{NS}
\end{IEEEeqnarraybox*}\right]=\\
(\mathbf{a}-\hat{\mathbf{a}})^{T}\Re{(\mathbf{K})}_{S},
\label{modification of gain work set selection}
\end{IEEEeqnarray}
where $\Re{(\mathbf{K})}_{S}\in \mathbb{R}^{N_{r}\times S}$ denotes the matrix constructed by all the columns corresponding to the work set $S$. Then we define intermediate variable vector $\Phi\in \mathbb{C}^{N_{r}}$, $\Re{(\Phi)}=\Re{(\mathbf{y})}-\Re{(\mathbf{K})}(\mathbf{a}-\hat{\mathbf{a}})$ and $\Im{(\Phi)}=\Im{(\mathbf{y})}-\Re{(\mathbf{K})}(\mathbf{b}-\hat{\mathbf{b}})$
Thus (\ref{gain work set selection1}) can be rewritten as 
\begin{IEEEeqnarray}[\relax]{l}
\bigtriangledown \theta_{S}^{r}=-\frac{1}{2}(\Delta_{S}-\hat{\Delta})^{T}_{S}\Re{(\mathbf{K})}_{SS}(\Delta-\hat{\Delta})_{S}+(\Re{(\Phi)}_{S})^{T}(\Delta-\hat{\Delta})_{S}-\epsilon\mathbf{e}^{T}_{S}(\Delta+\hat{\Delta})_{S}
\label{gain work set selection2}
\end{IEEEeqnarray} 
Because the offset term is omitted in Large MIMO regression model, therefore different from SMO type algorithms, there is no linear equation constraint, which is inducted by offset $b$, as shown in (\ref{partial4}). Therefore it is possible to update only one variable pair in each iteration. However, more efficient work set selection strategies based on maximum dual objective gain selection without considering offset term, is proposed in \cite{steinwart2011training}.   A work set of size two is updated in each iteration. The computational cost is reduced while maintaining the comparable performance with that with offset. In CSVR large MIMO detector, the work set selection strategy in \cite{steinwart2011training} is exploited. Basically, in each iteration this strategy uses sequential single variable searchings, whose searching time is $O(n)$, where $n$ is the number of data set, to determine a work set of size two. On the one hand, this strategy can find a work set of size two, whose update approximately maximize the gain in the dual objective function. On the other hand, the brute-force searching for the work size of two, whose update can maximize the gain in the dual objective function, requires $O(n^{2})$ searching time. The former strategy requires as few iterations as the latter one, since in each iteration, the latter strategy is more expensive, the former one can enjoy significantly faster running speed.   
\subsection{Single Direction Solver}\label{single direction solver}
Recall the KKT complementary conditions 
\begin{IEEEeqnarray}[\relax]{l}
\label{KKT complementary1}
(C\tilde{R}(\xi_{i}^{r})-\alpha_{i})\xi_{i}^{r}=0, i=1,2,\ldots, L\\
\label{KKT complementary2}
(C\tilde{R}(\hat{\xi}_{i}^{r})-\hat{\alpha}_{i})\hat{\xi}_{i}^{r}=0, i=1,2,\ldots, L\\
\label{KKT complementary3}
\alpha(\epsilon+\xi_{i}^{r}-\Re{(\mathbf{y}_{i})}+<\mathbf{h}_{i},\mathbf{w}>_{\mathbb{H}})=0, i=1,2,\ldots, L\\
\label{KKT complementary4}
\hat{\alpha}_{i}(\epsilon+\hat{\xi}_{i}^{r}-<\mathbf{h}_{i},\mathbf{w}>_{\mathbb{H}}+\Re{(\mathbf{y}_{i})})=0, i=1,2,\ldots, L\\\nonumber
\label{complementary KKT condition2}
\end{IEEEeqnarray}
notice that in (\ref{KKT complementary3}) and (\ref{KKT complementary4}), there is at least one of $\epsilon+\xi_{i}^{r}-\Re{(\mathbf{y}_{i})}+<\mathbf{h}_{i},\mathbf{w}>_{\mathbb{H}}$ and  $\epsilon+\hat{\xi}_{i}^{r}-<\mathbf{h}_{i},\mathbf{w}>_{\mathbb{H}}+\Re{(\mathbf{y}_{i})}$ can be non zero, therefore in order to satisfy the equalities, at least one of $\alpha_{i}$ and $\hat{\alpha}_{i}$ need to be zero, that is $\alpha_{i}\hat{\alpha}_{i}=0$, because $0\leq \alpha_{i}(\hat{\alpha}_{i})\leq C\tilde{R}(\xi_{i}^{r})(C\tilde{R}(\hat{\xi}_{i}^{r}))$ and $\alpha_{i}\hat{\alpha}_{i}=0$, there is at least one of $C\tilde{R}(\xi_{i}^{r})-\alpha_{i}$ and $C\tilde{R}(\hat{\xi}_{i}^{r})-\hat{\alpha}_{i}$ in (\ref{KKT complementary1}) and (\ref{KKT complementary2}) is non zero, in order to satisfy the equalities, there is at least one of $\xi_{i}$ and $\hat{\xi}_{i}$ need to be zero, that is $\xi_{i}^{r}\hat{\xi}_{i}^{r}=0$. Hence we can substitute $\lambda_{i}=\alpha_{i}-\hat{\alpha}_{i}$ and $|\lambda_{i}|=\alpha_{i}+\hat{\alpha}_{i}$, therefore the update unit is the single optimization variable $\lambda_{i}$ rather than the pair $\alpha_{i}$ and $\hat{\alpha}_{i}$.

  We first introduce 1-D work set selection strategy in which single optimization variable whose update can maximize the gain of dual objective function is chosen as the work set in each iteration. Define $\Lambda=[\lambda_{1}, \lambda_{2}, \ldots, \lambda_{N_{r}}]$, $\sigma_{i}=\lambda_{i}^{new}-\lambda_{i}$, $\Sigma=[\sigma_{1}, \sigma_{2}, \ldots, \sigma_{N_{r}}]$. Substitute $\mathbf{a}, \hat{\mathbf{a}}$ with $\Lambda$ and using $\Sigma$. Let $\Lambda_{S}$ and $\Sigma_{S}$ denote the vector that consist of the components in $\Lambda$ and $\Sigma$ that belong to work set $S$, the sub dual objective function in (\ref{subset optimization}) and its gain (\ref{gain work set selection2})
  \begin{equation}
\max_{\Lambda_{S}}\quad \theta_{S}^{r}=-\frac{1}{2}[\Lambda_{S}^{T}\Re{(\mathbf{K})}_{SS}\Lambda_{S}]+[\Re{(\mathbf{y})}_{S}^{T}-\Lambda_{N}^{T}\Re{(\mathbf{K})}_{NS}
]\Lambda_{S}-\epsilon<\mathbf{e}_{S}^{T}, |\Lambda_{S}|>,
\label{subset optimization lambda}
\end{equation}

\begin{IEEEeqnarray}[\relax]{l}
\bigtriangledown \theta_{S}^{r}=-\frac{1}{2}\Sigma^{T}_{S}\Re{(\mathbf{K})}_{SS}\Sigma_{S}+(\Phi^{r}_{S})^{T}\Sigma_{S}-\epsilon<\mathbf{e}^{T}_{S},|\Lambda^{new}_{S}|-|\Lambda_{S}|>,
\label{gain work set selection lambda}
\end{IEEEeqnarray} 
In 1-D solver, $|S|=1$, the sub dual objective function corresponding to $\lambda_{m}$, $m=1,2,\ldots, N_{r}$ can be written as
\begin{IEEEeqnarray}[\relax]{l}
 \max_{\lambda_{m}}\quad \theta_{m}^{r}=-\frac{1}{2}(\lambda_{m}^{new})^{2}\Re{(\mathbf{K})}_{mm}
+[\Re{(\mathbf{y}_{m})}-\sum_{j\neq m}^{N_{r}}\Re{(\mathbf{K})}_{mj}\lambda_{j}]\lambda_{m}^{new}-\epsilon(|\lambda^{new}_{m}|),
 \label{optimization function 1-D}
 \end{IEEEeqnarray}
 Based on the definition of $\Phi$ in (\ref{gain work set selection2}), $\Re{(\Phi)}_{i}=\Re{(\mathbf{y}_{i})}-\sum_{j=1}^{N_{r}}\lambda^{r}_{j}\Re{(\mathbf{K})}_{ij}$, similarly, as to dual variable $\lambda^{i}_{i}$, $\Re{(\Phi)}_{i}^{i}=Im(y_{i})-\sum_{j=1}^{N_{r}}\lambda^{i}_{j}\Re{(\mathbf{K})}_{ij}$. Thereafter, for sake of brevity, we use $\lambda_{i}$ instead of $\lambda_{i}^{r}$.
Take the partial derivative of $\theta_{m}^{r}$ respect to $\tilde{\lambda}^{new}_{m}$,  we have 
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
\frac{\partial \theta_{m}^{r}}{\partial \tilde{\lambda}^{new}_{m}}=-\tilde{\lambda}_{m}^{new}\Re{(\mathbf{K})}_{mm}+\Re{(\mathbf{y}_{m})}-\sum_{j\neq m}^{N_{r}}\lambda_{j}\Re{(\mathbf{K})}_{mj}-\epsilon(sgn(\tilde{\lambda}^{new}_{m}))=\\
\nonumber
-\tilde{\lambda}_{m}^{new}\Re{(\mathbf{K})}_{mm}+\Re{(\Phi)}_{m}+\lambda_{m}\Re{(\mathbf{K})}_{mm}-\epsilon(sgn(\tilde{\lambda}^{new}_{m}))\\
\Rightarrow \tilde{\lambda}_{m}^{new}=\lambda_{m}+\frac{\Re{(\Phi)}_{m}-\epsilon(sgn(\tilde{\lambda}^{new}_{m}))}{\Re{(\mathbf{K})}_{mm}},
\label{partial optimization function 1-D lambda}
\end{IEEEeqnarray}

 The update of $\alpha_{1}$ or $\hat{\alpha}_{1}$ is completed by clipping
\begin{IEEEeqnarray}[\relax]{l}
\lambda^{new}_{m}=[\tilde{\lambda}^{new}_{m}]^{CR^{'}(\xi_{m})}_{-CR^{'}(\xi_{m})}
\label{clipped new dual variable}
\end{IEEEeqnarray}

where $[ ]_{a}^{b}$ denotes clipping function 
\begin{IEEEeqnarray}[\relax]{l}
[x]_{a}^{b}=\left\{\begin{array}{ll}
a \quad if\quad  x\leq a\\
x \quad if\quad a<x<b\\
b \quad if\quad x\geq b\\
\end{array}\right.
\label{clipping function}
\end{IEEEeqnarray}
based on (\ref{gain work set selection lambda}), The maximum gain of objective function respect to the $m$th optimization variable $\lambda_{m}$ is 
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
\bigtriangledown \theta_{m}^{r}=\theta_{m}^{r}(\lambda_{m}+\sigma_{m})-\theta_{m}^{r}(\lambda_{m})\\
\nonumber
=-\frac{1}{2}\sigma_{m}^{2}\Re{(\mathbf{K})}_{mm}+\Re{(\Phi)}_{m}\sigma_{m}-\epsilon(|\lambda^{new}_{m}|-|\lambda_{m}|)\\
=\sigma_{m}[-\frac{1}{2}\sigma_{m}\Re{(\mathbf{K})}_{mm}+\Re{(\Phi)}_{m}]-\epsilon(|\lambda^{new}_{m}|-|\lambda_{m}|),
\label{gain of 1-D objective function}
\end{IEEEeqnarray}
In 1-D searching procedure, the optimization variable whose update has the maximum gain of sub dual objective function is chosen, assume the $k$th optimization variable is chosen, that is 
\begin{equation}
k=\arg\max_{(m=1,2,\ldots, N_{r})} \bigtriangledown \theta_{m}^{r},
\label{1-D direction}
\end{equation}
\subsection{Double Direction Solver}\label{double direction solver}
Although the omission of offset in the CSVR-MIMO detector makes 1-D solver possible, however, recent work in machine learning field shows training SVM without offset by 2-D solver with special work set selection strategies has more rapid training speed while the comparable performance is retained\cite{steinwart2011training}. The optimal 2-D solver uses the same principle as 1-D solver, $|S|=2$, assume the work set consists of the $m$th and $n$th optimization variables, that are $\Lambda_{S}=[\lambda_{m}, \lambda_{n}]$. Based on (\ref{gain work set selection lambda}), the sub dual objective function can be written as 
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
\max_{\lambda_{m}, 
\lambda_{n}}\quad \theta^{r}_{m,n}=-\frac{1}{2}[(\tilde{\lambda}^{new}_{m})^{2}\Re{(\mathbf{K})}_{mm}+(\tilde{\lambda}^{new}_{n})^{2}\Re{(\mathbf{K})}_{nn}+2\tilde{\lambda}^{new}_{m}\tilde{\lambda}^{new}_{n}\Re{(\mathbf{K})}_{mn}]-\\
\nonumber
\tilde{\lambda}^{new}_{m}\sum_{j\neq m, n}^{N_{r}}\lambda_{j}\Re{(\mathbf{K})}_{mj}-\tilde{\lambda}^{new}_{n}\sum_{j\neq m, n}^{N_{r}}\lambda_{j}\Re{(\mathbf{K})}_{nj}+\Re{(\mathbf{y}_{m})}\tilde{\lambda}^{new}_{1}+\Re{(\mathbf{y}_{n})}\tilde{\lambda}^{2}_{n}\\
-\epsilon(|\lambda^{new}_{m}|+|\tilde{\lambda}_{n}^{new}|), 
\label{2-D solver}
\end{IEEEeqnarray}
Based on (\ref{2-D solver}), the partial derivatives of $\theta^{r}_{m,n}$ with respect to $\tilde{\lambda}^{new}_{m}$ and $\tilde{\lambda}_{n}^{new}$ are 
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
\label{lambda1 partial}
\frac{\partial \theta^{r}_{m,n}}{\partial \tilde{\lambda}^{new}_{m}}=-\tilde{\lambda}^{new}_{m}\Re{(\mathbf{K})}_{mm}-\tilde{\lambda}^{new}_{n}\Re{(\mathbf{K})}_{mn}-\sum_{j\neq m, n}^{N_{r}}\lambda_{j}\Re{(\mathbf{K})}_{mj}+\Re{(\mathbf{y}_{m})}-\epsilon sgn(\tilde{\lambda}^{new}_{m})=\\
-\tilde{\lambda}_{m}^{new}\Re{(\mathbf{K})}_{mm}-\tilde{\lambda}^{new}_{n}\Re{(\mathbf{K})}_{mn}+\Re{(\Phi)}_{m}+\lambda_{m}\Re{\mathbf{K}}_{mm}+\lambda_{n}\Re{(\mathbf{K})}_{mn}-\epsilon sgn(\tilde{\lambda}^{new}_{m})=0\\
\nonumber
\label{lambda2 partial}
\frac{\partial \theta^{r}_{m,n}}{\partial \tilde{\lambda}^{new}_{n}}=-\tilde{\lambda}^{new}_{n}\Re{(\mathbf{K})}_{nn}-\tilde{\lambda}^{new}_{m}\Re{(\mathbf{K})}_{mn}-\sum_{j\neq m, n}^{N_{r}}\lambda_{j}\Re{(\mathbf{K})}_{nj}+\Re{(\mathbf{y}_{n})}-\epsilon sgn(\tilde{\lambda}^{new}_{n})=\\
-\tilde{\lambda}_{n}^{new}\Re{(\mathbf{K})}_{nn}-\tilde{\lambda}^{new}_{m}\Re{(\mathbf{K})}_{mn}+\Re{(\Phi)}_{n}+\lambda_{m}\Re{(\mathbf{K})}_{mn}+\lambda_{n}\Re{(\mathbf{K})}_{nn}-\epsilon sgn(\tilde{\lambda}^{new}_{n})=0\\
\nonumber
\label{partial derivative sample1}
\end{IEEEeqnarray}
where $\frac{\partial |x|}{\partial x}=sgn(x)$ denotes the sign of $x$. Based on (\ref{lambda1 partial}) and (\ref{lambda2 partial}) we have 
\begin{IEEEeqnarray}[\relax]{l}
\label{post partial lambda1}
(\tilde{\lambda}^{new}_{m}-\lambda_{n})\Re{(\mathbf{K})}_{mm}=\Re{(\Phi)}_{m}-\epsilon sgn(\tilde{\lambda}^{new}_{m})-(\tilde{\lambda}^{new}_{n}-\lambda_{n})\Re{(\mathbf{K})}_{mn}\\
\label{post partial lambda2}
(\tilde{\lambda}^{new}_{n}-\lambda_{n})\Re{(\mathbf{K})}_{nn}=\Re{(\Phi)}_{n}-\epsilon sgn(\tilde{\lambda}^{new}_{n})-(\tilde{\lambda}^{new}_{m}-\lambda_{m})\Re{(\mathbf{K})}_{mn}\\
\nonumber
\end{IEEEeqnarray}
hence based on (\ref{post partial lambda1}) and (\ref{post partial lambda2}), the update formula of $\tilde{\lambda}^{new}_{m}$ and $\tilde{\lambda}^{new}_{n}$ are
\begin{equation}
\tilde{\lambda}^{new}_{m}=\lambda_{m}+\frac{\Re{(\Phi)}_{m}\Re{(\mathbf{K})}_{nn}-\Re{(\Phi)}_{n}\Re{(\mathbf{K})}_{mn}-\epsilon[sgn(\tilde{\lambda}^{new}_{m})\Re{(\mathbf{K})}_{nn}-sgn(\tilde{\lambda}^{new}_{n})\Re{(\mathbf{K})}_{mn}]}{\Re{(\mathbf{K})}_{mm}\Re{(\mathbf{K})}_{nn}-(\Re{(\mathbf{K})}_{mn})^{2}}
\label{update lambda1}
\end{equation}

\begin{equation}
\tilde{\lambda}^{new}_{n}=\lambda_{n}+\frac{\Re{(\Phi)}_{n}\Re{(\mathbf{K})}_{mm}-\Re{(\Phi)}_{m}\Re{(\mathbf{K})}_{mn}-\epsilon[sgn(\tilde{\lambda}^{new}_{n})\Re{(\mathbf{K})}_{mm}-sgn(\tilde{\lambda}^{new}_{m})\Re{(\mathbf{K})}_{mn}]}{\Re{(\mathbf{K})}_{mm}\Re{(\mathbf{K})}_{nn}-(\Re{(\mathbf{K})}_{mn})^{2}}
\label{update lambda2}
\end{equation} 
Then the updated optimization variables are clipped by constraint 
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
\lambda_{i}^{new}=[\tilde{\lambda}_{i}^{new}]^{CR^{'}(\xi_{i})}_{-CR^{'}(\xi_{i})},\\
i=m,n
\label{cipped optimization variable}
\end{IEEEeqnarray}
Based on (\ref{gain work set selection2}),  the gain of 2-D solver objective function can be written as 
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
\bigtriangledown \theta^{r}_{mn}=-\frac{1}{2}[\sigma_{m}^{2}\Re{(\mathbf{K})}_{mm}+\sigma_{n}^{2}\Re{(\mathbf{K})}_{nn}+2\sigma_{m}\sigma_{n}\Re{(\mathbf{K})}_{mn}]+\Re{(\Phi)}_{m}\sigma_{m}+\Re{(\Phi)}_{n}\sigma_{n}\\
-\epsilon(|\lambda^{new}_{m}|-|\lambda_{m}|+|\lambda^{new}_{n}|-|\lambda_{n}|),
\label{gain of 2-D objective function1}
\end{IEEEeqnarray}
assume the $i$th and $j$th optimization variables are chosen, the optimization variables in 2-D solver have the same update rule as that of 1-D solver, that is 
\begin{equation}
[i,j]=\arg\max_{m,n=1,2,\ldots, N_{r}}\bigtriangledown \theta_{m,n}^{r}.
\label{dual optimization variables}
\end{equation}
Recall the gain of objective function of 1-D solver in (\ref{gain of 1-D objective function}), we obtain
 \begin{IEEEeqnarray}[\relax]{l}
\bigtriangledown \theta^{r}_{ij}=\bigtriangledown \theta^{r}_{i}+\bigtriangledown \theta^{r}_{j}-\sigma_{i}\sigma_{j}\mathbf{K}^{r}_{ij},
\label{gain of 2-D objective function1}
\end{IEEEeqnarray}
where $\bigtriangledown \theta^{r}_{i}$, $\bigtriangledown \theta^{r}_{j}$ denote gains of 1-D solver with $m$th and $n$th dual variable pairs are chosen. 
\subsection{Approximation to Optimal Double Direction Solver based on Single Direction Solver}\label{approximate 2D solver}
From (\ref{gain of 2-D objective function1}), it is obviously that the gain of 2-D solver is a summation of the gain of 2 independent 1-D solver and a correlation term $\sigma_{i}\sigma_{j}\mathbf{K}^{r}_{ij}$.

Obviously optimal 2-D coordinate combination ($i, j$) can be determined by comparing the gains of all the possibilities exhaustively, which requires $O(n^{2})$ times of searching. Based on (\ref{gain of 2-D objective function1}), we can approximate optimal 2-D solution by 1-D search approach, we will prove in large MIMO scenario, when $N_{t}$ is sufficient large, with channel hardening become effective, this approximation is very efficient. Here we propose two kinds of 1-D approximate searching strategy:
\subsubsection{ 1-D searching without damping}
do one round 1-D searching and calculate all the 1-D gain based on (\ref{gain of 1-D objective function}), then choose the coordinates with first and second largest 1-D gain as the candidates, then update the two candidates by 2-D solver as shown in (\ref{update lambda1}) and (\ref{update lambda2})
\subsubsection{ 1-D searching with damping}
do two rounds 1-D searchings, in the first round find optimization variable $i$ that can maximize 1-D gain, then in the second round, find $j$th optimization variable with the value of $i$th coordinate updated.

From (\ref{gain of 2-D objective function1}), it can be easily interpreted the efficient of 1-D approximation approach is majorly determined by the approximation ratio $\frac{\sigma_{i}\sigma_{j}\mathbf{K}^{r}_{ij}}{\bigtriangledown \theta^{r}_{i}+\bigtriangledown \theta^{r}_{j}}$, hence we provide theoretical analyse from the view of channel hardening phenomenon. Prior the theoretical analyse, we first investigate some mathematical properties of channel hardening (to be completed).

For 1-D solver the gradient of (\ref{partial optimization function 1-D lambda}) with respect to $\lambda$ can be written as
\begin{equation}
\lambda_{1}^{new}=\lambda_{1}+\frac{\Phi_{1}-sgn(\lambda_{1}^{new})\epsilon}{\mathbf{K}^{r}_{11}},
\label{gradient lambda}
\end{equation}
In the update process the $sgn(\lambda^{new})$ is unknown at current step, therefore, we need to consider both the case $sgn(\lambda^{new})=-1 $ or $1$ and choose the one with the larger objective function gain $\bigtriangledown \theta^{r}_{S}$.

%\begin{figure}[htb]
%\centering
%\includegraphics[width=0.7\textwidth, height=5cm]{logE_om_em.eps}
%\caption{Empirical Estimation $E[\ln(\phi_{om})]_{em}$}
%\label{figure1}
%\end{figure}

%\begin{figure}[htb]
%\centering
%\includegraphics[width=0.7\textwidth, height=5cm]{logE_om_t.eps}
%\caption{Theoretical $E[\ln(\phi_{om})]_{t}$}
%\label{figure2}
%\end{figure}

%\begin{figure}[htb]
%\centering
%\includegraphics[scale=0.6]{NtE.eps}
%\caption{Relation between $N_{t}$ and $E[\ln(\phi_{om})]_{t}$}
%\label{figure3}
%\end{figure}

\section{Stopping Criteria}\label{stopping criteria}
As we have explained in section \ref{section lagrange duality}, the upper bound of Lagrangian dual objective function is determined by primal objective function, further more the optimal of primal and dual objective function is found if and only if the equality holds, that is 
\begin{equation}
\theta(\lambda^{r}, \lambda^{i})=f(\mathbf{w},\xi)
\label{duality equal}
\end{equation}
\begin{equation}
\frac{1}{2}||\mathbf{w}||^{2}_{\mathbb{H}}+C\sum_{i=1}^{N_{r}}[R(\xi^{r}_{i})+R(\hat{\xi}^{r}_{i})+R(\xi^{i}_{i})+R(\hat{\xi}^{i}_{i})],
\label{simple primal}
\end{equation}
(\ref{final complex lagrange duality}) can be rewritten as follow by substituting $\lambda^{r}=\alpha-\hat{\alpha}$, $|\lambda^{r}|=\alpha+\hat{\alpha}$ and $\lambda^{i}=\beta-\hat{\beta}$, $|\lambda^{i}|=\beta+\hat{\beta}$
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
\theta(\lambda^{r}, \lambda^{i})=-\frac{1}{2}<(\lambda^{r})^{T}, \mathbf{K}^{r}\lambda^{r}>-\frac{1}{2}<(\lambda^{i})^{T}, \mathbf{K}^{r}\lambda^{i}>+<Re(\mathbf{y})^{T}, \lambda^{r}>+<Im(\mathbf{y})^{T}, \lambda^{i}>\\
-\epsilon<\mathbf{e}^{T}, (|\lambda^{r}|+|\lambda^{i}|)>+C\sum_{i=1}^{N_{r}}[\tilde{R}(\xi^{r}_{i})+\tilde{R}(\hat{\xi}^{r}_{i})+\tilde{R}(\xi^{i}_{i})+\tilde{R}(\hat{\xi}^{i}_{i})],
\label{simple dual function}
\end{IEEEeqnarray}
Similarly, (\ref{complex duality regularization1}) can be formulated as 
\begin{IEEEeqnarray}[\relax]{l}
||\mathbf{W}||^{2}_{\mathbb{H}}=<(\lambda^{r})^{T}, \mathbf{K}^{r}\lambda^{r}>+<(\lambda^{i})^{T}, \mathbf{K}^{r}\lambda^{i}>-2<\lambda^{r}, \mathbf{K}^{i}\lambda^{i}>,
\label{simple regularization}
\end{IEEEeqnarray}
hence, duality gap can be formulated as
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
G(\lambda^{r}, \lambda^{i})=f(\mathbf{w},\xi)-\theta(\lambda^{r},\lambda^{i})=<(\lambda^{r})^{T}, \mathbf{K}^{r}\lambda^{r}>+<(\lambda^{i})^{T}, \mathbf{K}^{r}\lambda^{i}>-<Re(\mathbf{y})^{T}, \lambda^{r}>-<Im(\mathbf{y})^{T}, \lambda^{i}>\\
-\epsilon<\mathbf{e}^{T}, (|\lambda^{r}|+|\lambda^{i}|)>+C\sum_{i=1}^{N_{r}}[\xi^{r}_{i}R^{'}(\xi^{r}_{i})+\hat{\xi}^{r}_{i}R^{'}(\hat{\xi}^{r}_{i})+\xi^{i}_{i}R^{'}(\xi^{i}_{i})+\hat{\xi}^{i}_{i}R^{'}(\hat{\xi}^{i}_{i})]-2<\lambda^{r}, \mathbf{K}^{i}\lambda^{i}>.
\label{simple duality gap}
\end{IEEEeqnarray}
As we explained in section \ref{risk functional}, the choice of risk function is determined by distribution of noise, as to Gaussian noise, the risk function is 
\begin{equation}
R(\xi)=\frac{1}{2}\xi^{2},
\label{risk function1}
\end{equation} 
hence 
\begin{equation}
\tilde{R}(\xi)=R(\xi)-\xi R{'}(\xi)=-\frac{1}{2}\xi^{2},
\label{risk function2}
\end{equation}
In $\epsilon$-SVR, the objective to employ slack variables $\xi$ is to deal with the outliers that outside $\epsilon$ tube to compensate the influence from noise.
Therefore 
\begin{IEEEeqnarray}[\relax]{l}
\xi^{r}_{i}=Re(\mathbf{y}_{i})-Re(<\mathbf{h}_{i},\mathbf{w}>_{\mathbb{H}})-\epsilon\\
\hat{\xi}^{r}_{i}=Re(<\mathbf{h}_{i},\mathbf{w}>_{\mathbb{H}})-Re(\mathbf{y}_{i})-\epsilon
\label{outlier1}
\end{IEEEeqnarray}
Because $\xi^{r}\hat{\xi}^{r}=0$ (estimation can only exceed $\epsilon$ tube in one direction), therefore there is only one of $\xi$ and $\hat{\xi}$ need to be considered, thus 
\begin{IEEEeqnarray}[\relax]{l}
\xi^{r}_{i}(\hat{\xi}^{r}_{i})=\max(0, |Re(\mathbf{y}_{i})-Re(<\mathbf{h}_{i}, \mathbf{w}>_{\mathbb{H}})|-\epsilon)\\
\xi^{i}_{i}(\hat{\xi}^{i}_{i})=\max(0, |Im(\mathbf{y}_{i})-Im(<\mathbf{h}_{i}, \mathbf{w}>_{\mathbb{H}})|-\epsilon)\\
\xi\hat{\xi}=0,
\label{outlier2}
\end{IEEEeqnarray} 
Therefore the risk function can be rewritten as 
\begin{IEEEeqnarray}[\relax]{l}
R(\xi_{i}^{r})+R(\hat{\xi}_{i}^{r})=\frac{1}{2}(|Re(\mathbf{y}_{i})-Re(<\mathbf{h}_{i}, \mathbf{w}>_{\mathbb{H}})|)^{2}_{\epsilon}\\
\label{simple risk function1}
R(\xi_{i}^{i})+R(\hat{\xi}_{i}^{i})=\frac{1}{2}(|Im(\mathbf{y}_{i})-Im(<\mathbf{h}_{i}, \mathbf{w}>_{\mathbb{H}})|)^{2}_{\epsilon}
\label{simple risk function2}
\end{IEEEeqnarray}
where $()_{\epsilon}$ denotes $\epsilon$ insensitive function as we mention in section \ref{risk functional}. Based on (\ref{complex duality regression function}), we have 
\begin{IEEEeqnarray}[\relax]{l}
Re(\mathbf{y}_{i})-Re(<\mathbf{h}_{i}, \mathbf{W}>_{\mathbb{H}})=Re(\mathbf{y}_{i})-\sum_{j=1}^{N_{r}}\lambda^{r}_{j}\mathbf{K}^{r}_{ij}+\sum_{j=1}^{N_{r}}\lambda^{i}_{j}\mathbf{K}^{i}_{ij}\\
Im(\mathbf{y}_{i})-Im(<\mathbf{h}_{i}, \mathbf{W}>_{\mathbb{H}})=Im(\mathbf{y}_{i})-\sum_{j=1}^{N_{r}}\lambda^{i}_{j}\mathbf{K}^{r}_{ij}-\sum_{j=1}^{N_{r}}\lambda^{r}_{j}\mathbf{K}^{i}_{ij}
\label{intermediate function1}
\end{IEEEeqnarray}
we define two intermediate variables $\Phi$ and $\Psi$
\begin{IEEEeqnarray}[\relax]{l}
\label{intermedia function 2a}
\Phi^{r}=Re(\mathbf{y})-\mathbf{K}^{r}
\lambda^{r};
\Phi^{i}=Im(\mathbf{y})-\mathbf{K}^{r}\lambda^{i}\\
\label{intermediate function 2b}
\Psi^{r}=\mathbf{K}^{i}\lambda^{i};
\Psi^{i}=-\mathbf{K}^{i}\lambda^{r}\\
\nonumber
\end{IEEEeqnarray} 
Therefore based on (\ref{simple risk function1})-(\ref{intermediate function 2b}), duality gap in (\ref{simple duality gap}) can be rewritten as 
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
G(\lambda^{r}, \lambda^{i})=<(\lambda^{r})^{T}, \mathbf{K}^{r}\lambda^{r}>+<(\lambda^{i})^{T}, \mathbf{K}^{r}\lambda^{i}>-<Re(\mathbf{y})^{T}, \lambda^{r}>-<Im(\mathbf{y})^{T}, \lambda^{i}>\\
+\epsilon<\mathbf{e}^{T}, (|\lambda^{r}|+|\lambda^{i}|)>+C\sum_{i=1}^{N_{r}}[(|\Phi^{r}_{i}+\Psi^{r}_{i}|)^{2}_{\epsilon}+(|\Phi^{i}_{i}+\Psi^{i}_{i}|)^{2}_{\epsilon}]-2<\lambda^{r}, \mathbf{K}^{i}\lambda^{i}>.
\label{simple duality gap final}
\end{IEEEeqnarray}
Based on objective function in (\ref{simple dual function}), (\ref{simple duality gap final}) can be rewritten as 
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
G=(\lambda_{r}, \lambda_{i})=<Re(\mathbf{y})^{T}, \lambda^{r}>+<Im(\mathbf{y})^{T}, \lambda^{i}>-\epsilon<\mathbf{e}^{T}, (|\lambda^{r}|+|\lambda^{i}|)>-2\theta(\lambda_{i}, \lambda_{j})\\
-2<\lambda^{r}, \mathbf{K}^{i}\lambda^{i}>.\label{simple duality gap ratio1}
\end{IEEEeqnarray}
The duality gap between primal problem and dual problem is used to evaluate how close a solution is to global minimum. In our scenario, duality gap is employed as stopping criteria.
Therefore to make stopping criteria more effective to monitor if algorithm convergent, we monitor the ratio by a value of tolerance (usually this tolerance is set to $10^{-3}$).
\begin{equation}
\frac{G}{G+\theta}
\label{simple duality gap ratio2}
\end{equation}
\subsection{Update $\Phi$, $\Psi$ and $G$}
In each iteration $\Phi$, $\Psi$ and $G$ are updated partially based on 2 updated optimization variables. Here we give the pseudo code to update $\Phi$, $\Psi$ and $G$.

Based on the definition of $\Phi$ and $\Psi$ in (\ref{intermedia function 2a}) and (\ref{intermediate function 2b}), we have the following procedure to update $\Phi$ and $\Psi$ in real channel and imaginary channel, assume the optimization coordinate updated in each channel are $1$ and $2$.\\
\begin{algorithm}[htb]
\begin{algorithmic}
\Procedure{\textbf{1}. Update $\Phi^{r}$ and $\Psi^{i}$ in real channel}{}
\For{$i=1:N_{r}$}
\State $\Phi_{i}^{r}=\Phi_{i}^{r}-\sigma_{1}^{r}\mathbf{K}^{r}_{i1}-\sigma_{2}^{r}\mathbf{K}_{i2}^{r}$
\State $\Psi_{i}^{i}=\Psi_{i}^{i}-\sigma_{1}^{r}\mathbf{K}^{i}_{i1}-\sigma_{2}^{r}\mathbf{K}^{i}_{i2}$
\EndFor
\EndProcedure
\end{algorithmic}
\end{algorithm}
\begin{algorithm}[htb]
\begin{algorithmic}
\Procedure{\textbf{2}. Update $\Phi^{i}$ and $\Psi^{r}$ in imaginary channel}{}
\For{$i=1:N_{r}$}
\State $\Phi_{i}^{i}=\Phi_{i}^{i}-\sigma_{1}^{i}\mathbf{K}^{r}_{i1}-\sigma_{2}^{i}\mathbf{K}_{i2}^{r}$
\State $\Psi_{i}^{r}=\Psi_{i}^{r}+\sigma_{1}^{i}\mathbf{K}^{i}_{i1}+\sigma_{2}^{i}\mathbf{K}^{i}_{i2}$
\EndFor
\EndProcedure
\end{algorithmic}
\end{algorithm}

Then the risk function term in (\ref{simple duality gap final}) is updated as following
\begin{algorithm}[htb]
\begin{algorithmic}
\Procedure{\textbf{3}. Update risk function in real channel}{$\chi^{r}$}
\State $\chi^{r}=0$  \Comment{initial risk term}
\For{$i=1:N_{r}$}
\If{$|\Phi^{r}_{i}+\Psi^{r}_{i}|>\epsilon$}
\State $\chi^{r}+=(|\Phi^{r}_{i}+\Psi^{r}_{i}|-\epsilon)^{2}$
\EndIf
\EndFor
\EndProcedure
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[htb]
\begin{algorithmic}
\Procedure{\textbf{4}. Update risk function in imaginary channel}{$\chi^{i}$}
\State $\chi^{i}=0$  \Comment{initial risk term}
\For{$i=1:N_{r}$}
\If{$|\Phi^{i}_{i}+\Psi^{i}_{i}|>\epsilon$}
\State $\chi^{i}+=(|\Phi^{i}_{i}+\Psi^{i}_{i}|-\epsilon)^{2}$
\EndIf
\EndFor
\EndProcedure
\end{algorithmic}
\end{algorithm}
The pseudo code to update duality gap $G$ based on (\ref{simple duality gap final}) is shown as follow assume the coordinate updated in real channel is $i$ and $j$, in imaginary channel is $m$ and $f$.
\begin{algorithm}[htb]
\begin{algorithmic}
\Procedure{\textbf{5}. Update $G$}{}
\State $G+=Re(\mathbf{y}_{1})\sigma^{r}_{i}+Re(\mathbf{y}_{2})\sigma^{r}_{j}$
\State $G+=Im(\mathbf{y}_{1})\sigma^{i}_{m}+Re(\mathbf{y}_{2})\sigma^{i}_{f}$
\State $G-=\epsilon(|\lambda^{r}_{i}+\sigma^{r}_{i}|-|\lambda^{r}_{i}|+|\lambda^{r}_{j}+\sigma^{r}_{j}|-|\lambda^{r}_{j}|)$
\State $G-=\epsilon(|\lambda^{i}_{m}+\sigma^{i}_{m}|-|\lambda^{i}_{m}|+|\lambda^{i}_{f}+\sigma^{i}_{f}|-|\lambda^{i}_{f}|)$
\State $G-=2(\bigtriangledown \theta_{i,j,m,f}(\sigma^{r}_{i},\sigma^{r}_{j},\sigma^{i}_{m},\sigma^{i}_{f}))$\Comment{Update sub objective function based on (\ref{gain of 2-D objective function1})}
\State $G+=C(\chi^{r}+\chi^{i})^{new}-(\chi^{r}+\chi^{i})$ \Comment{Update risk function term based on \textbf{Procedure 3} and \textbf{Procedure 4}}
\State $G-=\sigma^{r}_{i}\sigma^{i}_{m}\mathbf{K}^{i}_{im}+\sigma^{r}_{i}\sigma^{i}_{f}\mathbf{K}^{i}_{if}+\sigma^{r}_{j}\sigma^{i}_{m}\mathbf{K}^{i}_{jm}+\sigma^{r}_{j}\sigma^{i}_{f}\mathbf{K}^{i}_{jf}+\sigma^{i}_{m}\sum_{k=1}^{N_{r}}\lambda_{k}^{r}\mathbf{K}^{i}_{km}+\sigma^{i}_{f}\sum_{k=1}^{N_{r}}\lambda_{k}^{r}\mathbf{K}^{i}_{kf}-\sigma^{r}_{i}\sum_{k=1}^{N_{r}}\lambda_{k}^{i}\mathbf{K}^{i}_{ki}-\sigma^{r}_{j}\sum_{k=1}^{N_{r}}\lambda_{k}^{i}\mathbf{K}^{i}_{kj}$ \Comment{Update $<\lambda^{r},\mathbf{K}^{i}\lambda^{i}>$}
\EndProcedure
\label{update G Phi and Psi}
\end{algorithmic}
\end{algorithm}

Pseudo code for sequential single searching 2-D solver is shown as following 
\begin{algorithm}[htb]
\begin{algorithmic}
\Procedure{\textbf{6}. Sequential single searching 2-D solver without Damping}{} 
\State Step 1. Search for two optimization variables based on single direction solver
\For{$i=1:N_{r}$}
\State calculate $\bigtriangledown \theta^{r}_{i}(\bigtriangledown \theta_{i}^{i})$\Comment{Based on single direction solver \ref{single direction solver}}
\EndFor
\State choose the dual variable with first and the second largest gain of sub objective function, denoted as 1st and 2nd
\State Step 2. Update 1st and 2nd optimization variables based on double direction solver 
\State update $\lambda^{r}_{1st}(\lambda^{i}_{1st})$ and $\lambda^{r}_{2nd}(\lambda^{i}_{2nd})$\Comment{Based on double direction solver \ref{double direction solver}}
\State update $\Phi^{r}(\Phi^{i})$ and $\Psi^{r}(\Psi^{i})$ by \textbf{Procedure 1} and \textbf{Procedure 2}
\EndProcedure
\end{algorithmic}
%\caption{Sequential 2-D Solver without Damping}
\label{1D2D Nodamping}
\end{algorithm}

 \begin{algorithm}
\begin{algorithmic}
\Procedure{\textbf{7}. Sequential single searching 2-D solver with Damping}{}
\State Step 1. Search for two optimization variables based on single direction solver
\For{$i=1:N_{r}$}\Comment{First round searching}
\State calculate $\bigtriangledown \theta_{i}^{r}(\bigtriangledown \theta_{i}^{i})$ \Comment{Based on single direction solver \ref{single direction solver}}
\EndFor
\State choose the optimization variable with the largest gain of objective function as $1st_{1}$
\State update $\Phi^{r}(\Phi^{i})$ and $\Psi^{r}(\Psi^{i})$ with respect to $1st_{1}$
\For{$i=1:N_{r}$} \Comment{Second round searching}
\State calculate $\bigtriangledown \theta_{i}^{r}(\bigtriangledown \theta_{i}^{i})$ \Comment{Based on single direction solver \ref{single direction solver}}
\EndFor
\State choose the optimization variable with the largest gain of objective function as $1st_{2}$
\State Step 2. Update $1st_{1}$ and $1st_{2}$ optimization variables based on double direction solver 
\State update $\lambda^{r}_{1st_{1}}(\lambda^{i}_{1st_{1}})$ and $\lambda^{r}_{1st_{2}}(\lambda^{i}_{1st_{2}})$\Comment{Based on double direction solver \ref{double direction solver}}
\State update $\Phi^{r}(\Phi^{i})$ and $\Psi^{r}(\Psi^{i})$ by \textbf{Procedure 1} and \textbf{Procedure 2}

\EndProcedure
\end{algorithmic}
%\caption{Sequential 2-D Solver with Damping}
\label{1D2D damping}
\end{algorithm}

The following is the pseudo code of complex support vector detector (CSVD) is shown in Appendix \ref{Pseudo code of CSVD}

%\section{Hyperparameter Setting}
\newpage
\section{Computer Simulations}
Computer simulation is launched to test the detection and run time performance of proposed dual channel complex support vector detection algorithm. For sake of brevity, the real case is tested first, all the experiments are made by C, compiled by gcc version 4.8.3 on 64 bit Fedora (release 19) Linux system. The experiment platform is a desktop computer with I5-4th generation CPU with quad processing cores, 3.2 GHz clock rate, 8 GB RAM.

For sake of brevity, we consider a real uncoded spatial multiplex large MIMO system to simulate one channel of the proposed dual channel complex support vector detection algorithm.  with $N_{r}$ received antennas and $N_{t}$ transmitted antennas. The propagation channel matrix is constructed by channel gain components that are identically independent distributed (i.i.d) Gaussian random variables with zero mean and unit variance. transmitted symbols are mutually independent modulated by $M$ PAM with normalized average energy $\frac{1}{N_{t}}$, transmitted over flat fading channel, the sample of noise is AWGN with zero mean and variance $\frac{1}{10^{SNR/10}}$, where $SNR$ denotes the signal to noise ratio. 
We make experiment to low loading factor system $100\times 40$ and full loading factor $100\times 100$, with at least $1e^{5}$ channel realizations and at least $500$ symbol errors accumulated. Fig.\ref{SER performance} shows the symbol error rate (SER) performance, Table.\ref{iteration time} shows the average iteration time of real SVD for different SNR. 
\begin{figure}[htb]
\centering
%\def\svgscale{0.5}
\def\svgwidth{\columnwidth}
\input{BER_curve_real_system.pdf_tex}
\caption{SER performance of $100\times 100 $ and $100\times 40$ MIMO system}
\label{SER performance}
\end{figure}
\begin{table}[htb]
\renewcommand{\arraystretch}{1.3}
\caption{Average Iteration Time of Real Support Vector Detector}
\label{iteration time}
\centering
\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
\hline
\multirow{2}{*} {\bfseries Array Size} & \multicolumn{9}{|c|}{\bfseries SNR}\\
\cline{2-10}
 &6& 8 &10& 12& 14& 16& 18& 20& 22\\
\hline
$100\times 40$&682 &681 &681 &681& 680 &679 &678&677 & 680\\
\hline
$100\times 100$&1925 &1916& 1903& 1885& 1862& 1827& 1782&1723&1654 \\
\hline
\end{tabular}
\end{table}
% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table and, given that table
% captions serve much like titles, are usually capitalized except for words
% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
% and up, which are usually not capitalized unless they are the first or
% last word of the caption. Table text will default to \footnotesize as
% IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that the IEEE does not put floats in the very first column
% - or typically anywhere on the first page for that matter. Also,
% in-text middle ("here") positioning is typically not used, but it
% is allowed and encouraged for Computer Society conferences (but
% not Computer Society journals). Most IEEE journals/conferences use
% top floats exclusively. 
% Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the
% \fnbelowfloat command of the stfloats package.
%\section{Complexity Analysis}

%\section{Conclusion}
%The conclusion goes here.

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

\appendices
%\section{Proof of the First Zonklar Equation}
%Appendix one text goes here.

% you can choose not to have a title for an appendix
% if you want by leaving the argument blank
\section{Pseudo Code of CSVD}\label{Pseudo code of CSVD}
\begin{algorithm}[htb]
\begin{algorithmic}
\Procedure{CSVD}{$\mathbf{y}$,$\mathbf{H}$}
\State Step 1. Initialization
\State $\mathbf{K}=\mathbf{H}\mathbf{H}^{H}$ \Comment{kernel matrix}
\State $\chi^{r}=0$, $\chi^{i}=0$ \Comment{risk function}
\For{$i=1:N_{r}$}\Comment{initialize $\lambda^{r}$, $\lambda^{i}$, $\Phi^{r}$, $\Phi^{i}$, $\Psi^{r}$, $\Psi^{i}$ and duality gap $G$}
\State $\lambda_{i}^{r}=0, \lambda_{i}^{i}=0$
\State $\Phi_{i}^{r}=Re(y_{i}), \Phi_{i}^{i}=Im(y_{i})$
\State $\Psi_{i}^{r}=0, \Psi_{i}^{i}=0$
\If{$|\Phi^{r}_{i}|>\epsilon$}
\State $\chi^{r}+=(|\Phi^{r}_{i}|-\epsilon)^{2}$
\EndIf
\If{$|\Phi^{i}_{i}|>\epsilon$}
\State $\chi^{i}+=(|\Phi_{i}^{i}|-\epsilon)^{2}$
\EndIf
\EndFor 
\State $G=C(\chi^{r}+\chi^{i})$ \Comment{initialize duality gap}
\State $\theta=-0.5G$ \Comment{initialize objective function}
\State  Step 2. if $G>tol$, go to step 3, else go to Step 5
\State Step 3.
\State Sequentia single searching 2-D solver with or without damping \Comment{find two optimization variables to be updated} 
\State Step 4. \textbf{Procedure 5} update $G$
\State Step 5.
\State $\tilde{x}=(\lambda^{r}+i\lambda^{i})^{T}\mathbf{H}$ \Comment{reconstruct $\mathbf{x}$}
\State $\mathbf{x}=\mathbb{Q}(\tilde{x})$\Comment{$\mathbb{Q}(\cdot)$ denotes quantization operation based on symbol constellation}
\State go back to Step 2
\State Step 6. \textbf{Return} $\mathbf{x}$

\EndProcedure
\end{algorithmic}
\label{CSVD algorithm}
\caption{Dual Channel Complex Support Vector Detection Algorithm }
\end{algorithm}
%Appendix two text goes here.

% use section* for acknowledgment
%\section*{Acknowledgment}

%The authors would like to thank...

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
%\begin{thebibliography}{1}

%\bibitem{IEEEhowto:kopka}
%H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
%  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

%\end{thebibliography}

% biography section
% 
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

%\begin{IEEEbiography}{Michael Shell}
%Biography text here.
%\end{IEEEbiography}

% if you will not have a photo at all:
%\begin{IEEEbiographynophoto}{John Doe}
%Biography text here.
%\end{IEEEbiographynophoto}

% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

%\begin{IEEEbiographynophoto}{Jane Doe}
%Biography text here.
%\end{IEEEbiographynophoto}

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
%\end{spacing}

\newpage
\bibliographystyle{IEEEtran}
\bibliography{citation}	
\end{spacing}												
\end{document}