### 2015-11-09 Thesis Draft Tianpei Chen/Thesis Draft/C-ChannelHardening.tex

\resetdatestamp

\chapter{Theoretical Analysis of Channel Hardening Phenomenon}
\section{System Model}