### 2015-11-13 Thesis Draft Tianpei Chen/Thesis Draft/C-GA.tex

\resetdatestamp

\chapter{Low Complexity Near Optimal Hybrid Detector based on Genetic Algorithm}
In the previous chapter, we introduce two CSVPD aided local search based algorithms, and show the improvement CSVD-LAS -OPIC comparing to MMSE-LAS -OPIC.    
a current trend is the hybridization of of single solution based metaheuristics algorithms in population-based ones, working as intelligent subordinate heuristic operators. In this chapter, we introduce Hybrid Genetic Algorithms which utilizes OPIC as a genetic operator and accelerate the revolution process comparing to random revolution, in addition the performance of the proposed algorithm can be further improved by the hybridization with the principle from neighbour variable search. 
\section{Genetic Algorithm}