### 2015-11-18 Thesis Draft Tianpei Chen/Thesis Draft/C-GA.tex

\resetdatestamp

\chapter{Low Complexity Near Optimal Hybrid Detector based on Genetic Algorithm}\label{Chapter GA}
In the previous chapter, we introduce two CSVPD aided local search based algorithms, and show the improvement CSVD-LAS -OPIC comparing to MMSE-LAS -OPIC.    
a current trend is the hybridization of of single solution based metaheuristics algorithms in population-based ones, working as intelligent subordinate heuristic operators. In this chapter, we introduce Hybrid Genetic Algorithms which utilizes OPIC as a genetic operator and accelerate the revolution process comparing to random revolution. Furthermore, instead of randomly generated initial population, the solution of the CSVPD is fed to Hybrid GA and the initial generation consists of the neighbour of the initial solution from CSVPD.  
\section{CSVPD aided Hybrid Genetic Algorithm}
Genetic Algorithm (GA) is one of the population-based metaheuristic algorithms and can be viewed as a computational model of natural biological evolution process. GA works iteratively and in each iteration it handles a fixed size population of individuals. A number of intelligent genetic operators are used to update the population. Usually the individuals are a string of symbols that represent a candidate solution to the problems. GA has been applied in Code Division Multiple Access (CDMA) detection problem\cite{juntti1997genetic}\cite{ergun2000multiuser}.

Figure (\ref{}) shows the flow chart of the proposed hybrid GA. In the beginning, the initial solution of CSVPD is fed to GA, a initial population of solutions are generated based on this initial solution, then in each iteration, each candidate solution is evaluated and selected by fitness function. The elites are selected as parents which can generate offspring. The offsprings are generated by recombining a pair of parents, this process is called crossover. Then mutation operates to offsprings with a given possibility. OPIC is integrated in GA as an operator that can accelerate the process that GA approximate to optimal solution. The populations is finally reconstructed by the parents and their offsprings. The best solution is chosen when the stopping criteria satisfied.

\subsection{Initialization}
\subsection{Selection}
\subsection{Mutation}
\subsection{OPIC} 
\subsection{Stopping Criteria}