### 2015-11-18 Thesis Draft Tianpei Chen/Thesis Template/A-A.tex

\resetdatestamp

\chapter{\LaTeX{} Macros}
\label{A:LaTeXmacros}

The \LaTeX{} commands and macros used in formatting the title page for
 this document are shown in this appendix.

\section{Thesis Preamble}

The commands used to create the title page for a thesis are shown
 below.
The McGill University crest is brought in via a macro \verb|McGillCrest|
 which allows for setting the size and colour of an imported PostScript
 file which contains the actual crest.
The title page also includes a red separator line.