### 2015-11-18 Thesis Draft Tianpei Chen/Thesis Template/C-Intro.tex

\resetdatestamp

\chapter{Thesis Organization}

A thesis should present results in a scholarly fashion.
The following discusses the organization of a thesis, such as would be
 appropriate to presenting research results pertaining to Electrical \&
 Computer Engineering.

\section{Scope of a Thesis}

The terms of reference differ for a Master's Thesis and a Doctoral Thesis.
For the Master's Thesis, the Faculty of Graduate Studies and Research at
McGill University \cite{McGillTG:P1994} gives the following guidelines.

The terms of reference state that ``In most disciplines, Master's theses will
 not exceed 100 pages.''